### 2023-10-06 paper/main.tex
added approach section

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{fancyhdr}       % header
\usepackage{graphicx}       % graphics
\graphicspath{{media/}}     % organize your images and other figures under media/ folder

%Header
\pagestyle{fancy}
\thispagestyle{empty}
\rhead{ \textit{ }}

% Update your Headers here
\fancyhead[LO]{Running Title for Header}
% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}

%% Title
\title{Generating Synthetic Data via Intent-Preserving and Intent-Corrupting Augmentations for Training Dialogue Embeddings.
%%%% Cite as
%%%% Update your official citation here when published 
\thanks{\textit{\underline{Citation}}: 
\textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
}

\author{
  Author1, Author2 \\
  Affiliation \\
  Univ \\
  City\\
  \texttt{\{Author1, Author2\}email@email} \\
  %% examples of more authors
   \And
  Author3 \\
  Affiliation \\
  Univ \\
  City\\
  \texttt{email@email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}
Text embeddings from pre-trained language models have been proven to be extraordinarily useful for various sentence-level tasks, such as pair classification, similarity estimation, and retrieval. Corresponding models are usually trained on large amounts of clean and diverse data using contrastive loss. Unfortunately, there are no such datasets for the domain of dialogue data. In this work, we describe the process of mining a synthetic dataset of dialogues for contrastive learning with hard negatives. We investigate various augmentation strategies for constructing dialogues with preserved or corrupted intents (positive and negative samples, respectively). To demonstrate the stated cleanliness and diversity, we train a dialogue encoder model and analyze its properties.
\end{abstract}

% keywords can be removed
\keywords{text embedding \and dialogue \and synthetic data \and text augmentation}

\section{Introduction}
Obtaining embeddings is one of the key themes in machine learning as a whole, and a popular task in recent years. Vector representation of an object is a convenient mathematical entity. If an embedding accurately and comprehensively encodes the semantics of the original data, it opens up the possibility of using it in a wide range of tasks.

In the field of natural language processing, classical methods for text vectorization such as bag of words \cite{bow} and tf-idf \cite{SprckJones2021ASI} have long been invented. Thanks to deep learning, we have witnessed remarkable word vectorizations for their time, such as word2vec \cite{mikolov2013efficient} and GloVe \cite{pennington-etal-2014-glove}, which convey the semantic similarity between words; CoVe \cite{mccann2018learned} and ELMo \cite{peters-etal-2018-deep}, which provide information about the surrounding context. Recently, powerful encoder models have emerged that produce general-purpose text embeddings \cite{xiao2023cpack, wang2022text, li2023general}. They incorporate so much semantic information about texts that it can be applied to tasks such as classification, clustering, ranking, semantic textual similarity, and more. The success of these models is largely attributed to the use of contrastive learning on massive datasets.

However, the more specific the data structure, the more challenging it is to mine a large dataset. As of today, there are no encoding methods for dialogue data. In other words, there is no way to obtain a dense vector representation that conveys universal information about a dialogue. There are language models that adapt successfully to the hierarchical and temporal nature of dialogue \cite{zhang-etal-2023-dialog, li2022future}, but they only address tasks at the token and utterance levels, not at the dialogue level. This is due to the lack of datasets that pose challenges at the dialogue level.

Data is almost always scarce when it comes to building dialogue models. To date, numerous methods for generating synthetic dialogues have been devised \cite{kim2021neuralwoz, mohapatra2021simulated, wan-etal-2022-unified, zheng2023augesc}, but they do not generate dialogues in pairs \cite{schick2021generating}, as is conceptually important for contrastive embedding learning. The simplest way to expand a training dataset is through augmentation \cite{soudani2023data}. However, we find them insufficient, as they do not significantly alter the structure of the dialogue. In this work, we will describe a method for constructing a synthetic dataset of dialogue pairs using various augmentations that preserve or alter the set of intents in the dialogue. These augmentations significantly affect its structure.

\section{Approach}

\subsection{Augmentations}

\textbf{Token Insertion.} One of the simple yet effective ways to augment text is to lengthen it by inserting additional tokens. For this purpose, we added a special token `<mask>` to random places in the dialogues and used transformer models trained on the MLM task to fill these masks. Insertion was rejected if the token proposed by the model was only a fragment of a word (BPE, Wordpiece) or if the prediction probability was below a manually set threshold. To take the dialogue context into account during token insertion, multiple consecutive responses or the entire dialogue were fed into the MLM model at once.

\textbf{Token Replacement.} This method is identical to the previous one, except that instead of adding the `<mask>` token between tokens in the original dialogue, some tokens in the original dialogue are replaced.

\textbf{Back Translation.} Translation from the original language to another language and then back to the original language. MT models were used for this purpose.

\textbf{Shuffling Responses.} Previous augmentations modify the dialogue within individual phrases since they are methods applicable to arbitrary text data. It seems essential to learn how to change the order of responses in a dialogue to create a valid example of meaningful dialogue. For this purpose, we propose using models that measure the similarity between responses within a dialogue \ref{fig:pairwise}. Using these distances, it is possible to cluster responses within each dialogue. Experiments showed that these clusters represent significant individual stages of the dialogue that can be shuffled with each other.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.7\linewidth]{figures/dialogue-shufflers-pairwise.drawio.pdf}
    \caption{Model for measuring the distance between responses in a dialogue.}
    \label{fig:pairwise}
\end{figure}

\textbf{Shortening the Dialogue.} Individual clusters of responses within the dialogue can be discarded, resulting in a dialogue with fewer responses.

\textbf{Lengthening the Dialogue.} A model was trained to arrange responses in a dialogue based on a given list of responses. This transformer takes the text embeddings of each response as input without specifying information about the position of the response in the original dialogue. It outputs ranks or probabilities that can be used to "sort" the responses. If some external responses are added to the original dialogue, this model generates a new, longer dialogue.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.7\linewidth]{figures/dialogue-shufflers-listwise-lm-1.drawio.pdf}
    \caption{Model for extending the dialogue through sorting.}
    \label{fig:listwise}
\end{figure}

\textbf{Preserving and Altering Intent.} The augmentations mentioned above can be configured to either preserve or alter intent. As augmentations that preserve intent, we used all operations except token replacement, which was used to select hard negative examples.

\subsection{Dialogue Encoder}

\textbf{BERT.} As a baseline method for vectorization, we use BERT without any modifications. We input the dialogue as \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3 [SEP]}, and the output is the averaged vectors of the hidden state of the `SEP` tokens.

\textbf{HSSA and BiDeN.} For slightly advanced dialogue language models, we use HSSA and BiDeN.

For training dialogue encoders, we use contrastive learning with in-batch negative examples and hard negative examples:
$$
\mathcal{L}=-\log{\exp(\cos(x_i,y_i))\over\sum_{j=1}^B\exp(\cos(x_i,y_j))},
$$

\section*{Acknowledgments}
This was was supported in part by......

%Bibliography
\bibliographystyle{unsrt}  
\bibliography{references}

\end{document}

### 2023-10-22 paper/main.tex
changed figures
added appendix

@@
 \usepackage{graphicx}       % graphics
-\graphicspath{{media/}}     % organize your images and other figures under media/ folder
+\graphicspath{{figures/}}     % organize your images and other figures under figures/ folder
 
+\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools,esint,eucal}  %math
+\usepackage{subcaption}
 %Header
@@
 %% Title
-\title{Generating Synthetic Data via Intent-Preserving and Intent-Corrupting Augmentations for Training Dialogue Embeddings.
+\title{Generating Synthetic Data via Intent-Preserving and Intent-Corrupting Augmentations for Multi-task Training Dialogue Embeddings.
 %%%% Cite as
@@
 \author{
-  Author1, Author2 \\
-  Affiliation \\
-  Univ \\
-  City\\
-  \texttt{\{Author1, Author2\}email@email} \\
+  Ilya Alekseev \\
+  Neural Networks and Deep Learning Lab \\
+  Moscow Institute of Physics and Technology \\
+  \texttt{ilya\_alekseev\_2016@list.ru} \\
   %% examples of more authors
    \And
-  Author3 \\
-  Affiliation \\
-  Univ \\
-  City\\
-  \texttt{email@email} \\
+  Denis Kuznetsov \\
+  Neural Networks and Deep Learning Lab \\
+  Moscow Institute of Physics and Technology \\
+  \texttt{kuznetsov.dp@phystech.edu} \\
   %% \AND
@@
 % keywords can be removed
-\keywords{text embedding \and dialogue \and synthetic data \and text augmentation}
+\keywords{text embedding \and dialogue \and synthetic data \and augmentation \and natural language processing}
 
 \section{Introduction}
-Obtaining embeddings is one of the key themes in machine learning as a whole, and a popular task in recent years.
-Vector representation of an object is a convenient mathematical entity.
+Obtaining embeddings is one of the key tasks in machine learning and popular one in recent years.
+Vector representation of an object is a convenient mathematical object.
 If an embedding accurately and comprehensively encodes the semantics of the original data, it opens up the possibility of using it in a wide range of tasks.
 
-In the field of natural language processing, classical methods for text vectorization such as bag of words \cite{bow} and tf-idf \cite{SprckJones2021ASI} have long been invented.
-Thanks to deep learning, we have witnessed remarkable word vectorizations for their time, such as word2vec \cite{mikolov2013efficient} and GloVe \cite{pennington-etal-2014-glove}, which convey the semantic similarity between words; CoVe \cite{mccann2018learned} and ELMo \cite{peters-etal-2018-deep}, which provide information about the surrounding context.
+In the field of natural language processing, classical methods for text vectorization such as bag of words \cite{bow} and tf-idf \cite{SprckJones2021ASI} have long been discovered.
+Thanks to deep learning, we have witnessed remarkable word vectorizations such as word2vec \cite{mikolov2013efficient} and GloVe \cite{pennington-etal-2014-glove}, which convey the semantic similarity between words; CoVe \cite{mccann2018learned} and ELMo \cite{peters-etal-2018-deep}, which encode information about the surrounding context.
 Recently, powerful encoder models have emerged that produce general-purpose text embeddings \cite{xiao2023cpack, wang2022text, li2023general}.
@@
 However, the more specific the data structure, the more challenging it is to mine a large dataset.
-As of today, there are no encoding methods for dialogue data.
+As of today, there are no encoding methods for entire dialogues.
 In other words, there is no way to obtain a dense vector representation that conveys universal information about a dialogue.
 There are language models that adapt successfully to the hierarchical and temporal nature of dialogue \cite{zhang-etal-2023-dialog, li2022future}, but they only address tasks at the token and utterance levels, not at the dialogue level.
-This is due to the lack of datasets that pose challenges at the dialogue level.
+There are text encoders for utterances in a dialogue \cite{henderson2020convert, Zhou2022}, but not for the entire dialogue.
 
 Data is almost always scarce when it comes to building dialogue models.
-To date, numerous methods for generating synthetic dialogues have been devised \cite{kim2021neuralwoz, mohapatra2021simulated, wan-etal-2022-unified, zheng2023augesc}, but they do not generate dialogues in pairs \cite{schick2021generating}, as is conceptually important for contrastive embedding learning.
+To date, numerous methods for generating synthetic dialogues have been devised \cite{kim2021neuralwoz, mohapatra2021simulated, wan-etal-2022-unified, zheng2023augesc}, but they do not generate dialogues in pairs \cite{schick2021generating}, as is conceptually important for contrastive learning.
 The simplest way to expand a training dataset is through augmentation \cite{soudani2023data}.
@@
 In this work, we will describe a method for constructing a synthetic dataset of dialogue pairs using various augmentations that preserve or alter the set of intents in the dialogue.
-These augmentations significantly affect its structure.
+These augmentations can be used for contrastive learning together with other tasks for pretraining powerful dialogue embeddings.
+
+\section{Problem Formulation}
+
+\textbf{Dialogue Data.} A dialogue is defined as the following list:
+$$
+d=[(u_1, s_1), \ldots, (u_n, s_n)],
+$$
+where $u_i$ represents the utterance of a participant in the dialogue at step $s_i$.
+We are interested in so-called task-specific dialogues with two participants: the system and the user.
+With some approximation, they can be described as dialogues between a customer and a service worker (or a robot).
+During the dialogue, the customer has various intents that the worker strives to fulfill.
+These intents can be finding a restaurant and booking a table, calling a taxi, or purchasing a train ticket.
+We will consider two dialogues similar if they have a similar set of intents.
+
+\textbf{Augmentation.} By augmentation, we mean the generation of new valid examples by transforming existing ones.
+Valid examples are those that sufficiently resemble real-world data.
+Let $D$ be the set of valid objects.
+Then augmentation is a non-identical mapping $\text{aug}(d)$ that does not take objects outside the set of valid objects:
+$$
+\text{aug}: D\to D.
+$$
+Typically, this generation is implemented by making small changes to a valid training object.
+For example, image augmentation might involve slight rotations or blurring.
+Text augmentation is especially challenging because validity implies adherence to language rules, meaningfulness and a certain style.
+In the case of dialogues, it is necessary to maintain the structure and role differentiation, as indicated earlier.
+
+\textbf{Embedding.} Embedding is a mapping of $D$ to a vector space:
+$$
+e:D\to E\subseteq\mathbb{R}^d.
+$$
+For an object $d$, its embedding $e(d)$ should convey some semantics of $d$.
+This is reflected in the fact that $e(d)$ may contain lexical information or latent features useful for classification and other downstream tasks.
+It is especially valuable if using embeddings $e(a), e(b)$ allows for assessing the semantic similarity of objects $a, b$.
+
+\section{Related Works}
+
+\texttt{this section is incomple, this is still a draft }
+
+AugSBERT \cite{thakur2021augmented} can be viewed as the similar approach to ours one, but their augmentation uses an already presented dataset of text pairs, which is absent in our case.
 
@@
 
-\subsection{Augmentations}
+\subsection{Augmentations} \label{sec:aug}
 
-\textbf{Token Insertion.} One of the simple yet effective ways to augment text is to lengthen it by inserting additional tokens.
-For this purpose, we added a special token `<mask>` to random places in the dialogues and used transformer models trained on the MLM task to fill these masks.
-Insertion was rejected if the token proposed by the model was only a fragment of a word (BPE, Wordpiece) or if the prediction probability was below a manually set threshold.
-To take the dialogue context into account during token insertion, multiple consecutive responses or the entire dialogue were fed into the MLM model at once.
+\textbf{Token Insertion.} One of the simple yet effective ways to augment text is to lengthen it by inserting extra tokens.
+For this purpose, we added a special token `<mask>` to random places in the dialogues and used transformer models trained on the MLM task to fill these masks \cite{devlin2019bert}.
+Insertion is rejected if the token proposed by the model is only a fragment of a word \cite{wu2016googles, sennrich-etal-2016-neural} or if the prediction probability is below a manually set threshold.
+To take the dialogue context into account during token insertion, multiple consecutive responses were fed into the mask-filling model at once as a compromise between feeding single utterances and entire dialogue.
 
-\textbf{Token Replacement.} This method is identical to the previous one, except that instead of adding the `<mask>` token between tokens in the original dialogue, some tokens in the original dialogue are replaced.
+\textbf{Token Replacement.} This method is identical to the previous one, except that instead of adding the `<mask>` token, some tokens in the original dialogue are replaced.
+In this case, the mask-filling model is fed with single utterances to make replacements more diverse and random.
 
-\textbf{Back Translation.} Translation from the original language to another language and then back to the original language. MT models were used for this purpose.
+\textbf{Back Translation.} Translation from the original language to another language and then back to the original language.
+Neural machine translation models were used for this purpose \cite{TiedemannThottingal}.
 
-\textbf{Shuffling Responses.} Previous augmentations modify the dialogue within individual phrases since they are methods applicable to arbitrary text data.
-It seems essential to learn how to change the order of responses in a dialogue to create a valid example of meaningful dialogue.
-For this purpose, we propose using models that measure the similarity between responses within a dialogue \ref{fig:pairwise}.
-Using these distances, it is possible to cluster responses within each dialogue.
+\textbf{Shuffling Utterances.} Previous augmentations modify the dialogue within a single utterance, since they are methods applicable to arbitrary text data.
+It seems essential to learn how to change the order of utterances in a dialogue to create new valid dialogue.
+For this purpose, we propose using a model that measures the similarity between utterances within a dialogue.
+Using these similarities, it is possible to cluster utterances within each dialogue.
 Experiments showed that these clusters represent significant individual stages of the dialogue that can be shuffled with each other.
 
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.7\linewidth]{figures/dialogue-shufflers-pairwise.drawio.pdf}
-    \caption{Model for measuring the distance between responses in a dialogue.}
-    \label{fig:pairwise}
-\end{figure}
+\textbf{Shortening Dialogue.} Individual clusters of utterances within the dialogue can be discarded, resulting in a pruned dialogue with fewer utterances.
 
-\textbf{Shortening the Dialogue.} Individual clusters of responses within the dialogue can be discarded, resulting in a dialogue with fewer responses.
-
-\textbf{Lengthening the Dialogue.} A model was trained to arrange responses in a dialogue based on a given list of responses.
-This transformer takes the text embeddings of each response as input without specifying information about the position of the response in the original dialogue.
-It outputs ranks or probabilities that can be used to "sort" the responses.
+\textbf{Lengthening Dialogue.} The special model was trained to arrange a list of given utterances.
+This transformer takes the text embeddings of each utterance as input sequence, without specifying information about their order in the original dialogue.
+It outputs ranks that can be used to "sort" the utterances to restore the original order.
 If some external responses are added to the original dialogue, this model generates a new, longer dialogue.
 
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.7\linewidth]{figures/dialogue-shufflers-listwise-lm-1.drawio.pdf}
-    \caption{Model for extending the dialogue through sorting.}
-    \label{fig:listwise}
-\end{figure}
+The augmentations mentioned above can be configured to either preserve or alter intent.
+Specifically, token replacement can be viewed as intent-corrupting augmentation, because all the keywords such as "restaurant", "taxi" etc.
+tend to be replaced.
+Pruning dialogue may remove some intents, but the result is still much similar to original dialogue, since its intents are fully encompassed by the original dialogue's intents.
+Shuffling utterances doesn't change any intents, it only changes their order.
+The rest of augmentations preserves intents because they either perform paraphrasing (back translation), or add new information (token insertion, lengthening dialogue).
 
-\textbf{Preserving and Altering Intent.} The augmentations mentioned above can be configured to either preserve or alter intent.
-As augmentations that preserve intent, we used all operations except token replacement, which was used to select hard negative examples.
+To expand the set of augmentations even further, we use a composition of several. More details about augmentation implementations are presented in appendix.
 
-\subsection{Dialogue Encoder}
+\subsection{Dialogue Encoder Architecture}
 
-\textbf{BERT.} As a baseline method for vectorization, we use BERT without any modifications.
-We input the dialogue as \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3 [SEP]}, and the output is the averaged vectors of the hidden state of the `SEP` tokens.
+As a baseline method for vectorization, we use RoBERTa \cite{liu2019roberta} without any modifications.
+The input is \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3 [SEP]}, and the output is the hidden state of \texttt{[CLS]} token from the last layer.
 
-\textbf{HSSA and BiDeN.} For slightly advanced dialogue language models, we use HSSA and BiDeN.
+For a slightly advanced dialogue language model, we use HSSA \cite{zhang-etal-2023-dialog}.
+It uses BERT \cite{devlin2019bert} as a backbone and modifies its attention to capture the hierarchical structure of dialogue and reach the computational trade-off between feeding a transformer separate utterances and feeding an entire dialogue.
 
-For training dialogue encoders, we use contrastive learning with in-batch negative examples and hard negative examples:
+\subsection{Pre-train Tasks}
+
+To make the encoder produce rich vector representations, it is necessary to train it to perform tasks where semantic features are engaged. A popular task is contrastive learning:
 $$
-\mathcal{L}=-\log{\exp(\cos(x_i,y_i))\over\sum_{j=1}^B\exp(\cos(x_i,y_j))},
+\mathcal{L}=-\log{\exp(\text{sim}(x,y))\over\sum_{z}\exp(\text{sim}(x,z))}.
 $$
+Here, $x=e_\theta(a), y=e_\theta(b)$ are embeddings of a pair of semantically similar objects and $z=e_\theta(c)$ is semantically distant from $x$, $\text{sim}$ is a metric similarity function.
+This task trains embeddings to convey semantic similarity as metric similarity.
 
@@
 
+\appendix
+\section{Dialogue Dataset}
+Large dataset is important for contrastive learning.
+For training our models, we took a merge of some task-oriented datasets from DialogStudio collection \cite{zhang2023dialogstudio}.
+All of them are listed in the table \ref{tab:dataset}.
+
+\begin{table}[!htb]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        Name & \# dialogues & \# utterances & \# tokens \\
+        \hline
+        AirDialogue & 321K & 4086K & 49.4M \\
+        SimJointGEN & 100K & 1584K & 22.5M \\
+        MS-DC & & & \\
+        MetaLWOZ & & & \\
+        MULTIWOZ2\_2 & & & \\
+        SGD & & & \\
+        KETOD & & & \\
+        FRAMES & & & \\
+        Disambiguation & & & \\
+        ABCD & & & \\
+        AirDialogue & & & \\
+        BiTOD & & & \\
+        Taskmaster1 & & & \\
+        \textbf{Total} & & & \\
+        \textbf{Filtered} & 501K & 6320K & 83.7M
+    \end{tabular}
+    \caption{All the datasets are taken from DialogStudio collection \cite{zhang2023dialogstudio}}
+    \label{tab:dataset}
+\end{table}
+
+\section{Auxiliary Models} \label{app:aug}
+
+As mentioned in Section \ref{sec:aug}, we trained special models to perform dialogue-level augmentations.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.6\linewidth}
+        \centering
+        \includegraphics[width=0.9\linewidth]{figures/pairwise.drawio.pdf}
+        \caption{for measuring similarities between utterances}
+        \label{fig:pairwise}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.35\linewidth}
+        \centering
+        \includegraphics[width=0.9\linewidth]{figures/listwise.drawio.pdf}
+        \caption{for lengthening dialogue}
+        \label{fig:listwise}
+    \end{subfigure}
+    \caption{Auxiliary models for performing augmentations.}
+    \label{fig:enter-label}
+\end{figure}
+
+\subsection{Pairwise Model}
+
+We utilized a model for measuring the similarity between dialog utterances.
+In its basic form, it can be implemented as follows: take sentence embeddings of the utterances and compare the cosine similarities between them.
+Training such a model on sequential utterances using contrastive learning yields commendable utterance embeddings \cite{Zhou2022}.
+
+However, this approach has a significant drawback; it does not consider the context from several preceding utterances.
+In a dialogue, it is crucial to compare not just pairs of utterances but pairs of context-response.
+Therefore, we employed the following model (fig.
+\ref{fig:pairwise}):
+
+\begin{enumerate}
+    \item  First, embeddings for all context utterances $c=[u_1,\ldots,u_k]$ and the response $r$ are obtained using a pretrained sentence encoder.
+    \item The embeddings of the context are concatenated and passed to a projector that outputs a vector representing the context.
+    \item The response embedding is fed into a second encoder, resulting in a vector representing the response.
+    \item The cosine similarity between the obtained vectors is computed as a measure of the context and response similarity.
+\end{enumerate}
+
+\texttt{aws-ai/dse-bert-large} model from hugging face \cite{wolf2020huggingfaces} was used as the sentence encoder.
+The model was trained with a contrastive loss using in-batch negative sampling, with the following parameters: batch size is 128, temperature is 0.05, context size is 3, projection size is 256.
+Only 3 last layers of sentence encoder were fine-tuned in order to decrease computational cost.
+Trained model reaches 0.955 retrieval accuracy@5.
+
+Batches were formed from "context-response" pairs from the entire dialog dataset, where negative examples were not samples from the same dialog but entirely random examples from the dataset.
+This allows batching of arbitrary sizes, not limited to the dialog size, making the pre-training task more challenging.
+
+The resulting model closely resembles the ConveRT model \cite{henderson2020convert} for obtaining utterance embeddings.
+The drawbacks of the latter model are, firstly, that it is proprietary, and secondly, its architecture is highly specific and does not utilize the familiar BERT-like backbone.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/pairwise-cluster-heatmap.jpg}
+    \caption{All similarities between contexts and responses within a dialogue. It is easy to see, that consecutive utterances form clusters.}
+    \label{fig:pairwise-clister-heatmap}
+\end{figure}
+
+As a result, the obtained model is capable of recognizing individual stages in dialogues (fig.
+\ref{fig:pairwise-clister-heatmap}).
+This behavior is achieved due to two factors.
+Firstly, when a dialogue furthers to a new topic, the similarity between two consecutive utterances drops substantially.
+This is clearly visible, for example, in dialogues in which ordering a taxi replaces booking a table in a restaurant.
+Secondly, within one topic, there are also small drops in similarity in cases where there is a transition from one question to another.
+For example, the question “how many people should I book for?” replaces the question “which restaurant do you prefer?”.
+Moreover, since these questions relate to one topic, they remain close to themselves and distant to questions on other topics.
+Therefore, clusters are obtained.
+
+(!provide the dialogue and change pic to vector instead of raster!)
+
+\subsection{Listwise Model}
+
+We trained the special model to merge utterances of two different dialogues (fig.
+\ref{fig:listwise}).
+It is a transformer over text embeddings of the utterances.
+Sentence encoder is \texttt{sentence-transformers/all-mpnet-base-v2} from hugging face.
+We used 4-layer transformer with 4 attention heads and hidden dimension twice smaller than sentence encoder's one, i.e.
+384.
+Only 3 last layers of sentence encoder were fine-tuned in order to decrease computational costs.
+
+Output ranks are transformed with softmax function.
+Then KL-divergence between output and target probabilities are minimized.
+Target probabilities are defined as softmax over true ranks of utterances, i.e.
+$-i$ for $i$-th utterance in dialogue.
+
+Resulting model trains to "sort" given utterances.
+Thanks to the attention mechanism of transformers, this can be viewed as asking the children at physical education class to look at each other and line up by height.
+
+To measure the sorting quality, we need to utilize appropriate metric.
+All traditional ranking metrics such as nDCG are designed to compare with gold ranks, not just sorting quality.
+So during validation of our model, we were converting the ranks to a permutation over the original sequence of $n$ elements.
+Then, we calculated the number of transpositions.
+It is easy to implement and can normalized by maximum possible number of transpositions $n(n-1)/2$, resulting in $[0,1]$-ranged metric.
+Our trained model reaches $0.96$ value.
+
+\section{Composition of Augmentations}
+
+In order to maximize diversity of training data, we use not only 5 basic augmentations described in section \ref{sec:aug}, but also 4 extra compositions of augmentations.
+All the resulting pipelines are defined in fig.
+\ref{fig:aug-compositions}.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.5\linewidth}
+        \includegraphics[width=0.85\linewidth]{figures/augmentation-pipeline-positive.drawio.pdf}
+        \caption{intent-preserving}
+        \centering
+    \end{subfigure}
+    \begin{subfigure}[t]{0.35\linewidth}
+        \includegraphics[width=0.9\linewidth]{figures/augmentation-pipeline-negative.drawio.pdf}
+        \caption{intent-corrupting}
+        \centering
+    \end{subfigure}
+    \caption{Compositions of augmentations.}
+    \label{fig:aug-compositions}
+\end{figure}
+
 \end{document}

### 2023-10-22 src/mylib/dialogue_encoder/filter_dataset_by_length.py
added letter C

@@ -1,0 +1,68 @@
+import json
+import os
+from transformers import AutoTokenizer
+from tqdm import tqdm
+import sys
+sys.path.insert(0, '/home/alekseev_ilya/dialogue-augmentation/nup')
+from models import SimpleDialogueEncoder
+
+
+def is_short_enough(dia, tokenizer, upper_bound=512):
+    """dia should be shorter than 512 minus number of SEP and CLS tokens"""
+    input_ids = SimpleDialogueEncoder._tokenize(tokenizer, [dia])['input_ids'][0]
+    return len(input_ids) <= upper_bound
+
+
+def filter_dataset(path_in, path_out, tokenizer='roberta-base', name_validator=None, number_extractor=None):
+    """copies all json chunks of a dataset from `path_in` to `path_out`. Each dia exceeding the length limit is replaced with None."""
+
+    if name_validator is None:
+        name_validator = lambda x: x.endswith('.json') and not x.startswith('ru')
+    
+    if number_extractor is None:
+        number_extractor = lambda x: int(x.split('.')[0])
+
+    tokenizer = AutoTokenizer.from_pretrained(tokenizer)
+        
+    if not os.path.exists(path_out):
+        os.makedirs(path_out)
+    chunk_names = [filename for filename in os.listdir(path_in) if name_validator(filename)]
+    chunk_names = sorted(chunk_names, key=number_extractor)
+    
+    for chunk_name in tqdm(chunk_names):
+        chunk_path_in = os.path.join(path_in, chunk_name)
+        chunk = json.load(open(chunk_path_in, 'r'))
+        
+        filtered_chunk = []
+        for i, dia in enumerate(chunk):
+            res = dia
+            if dia is not None and not is_short_enough(dia, tokenizer):
+                res = None
+                print(f'rejected dia #{i}')
+            filtered_chunk.append(res)
+        
+        chunk_path_out = os.path.join(path_out, chunk_name)
+        json.dump(filtered_chunk, open(chunk_path_out, 'w'))
+
+
+if __name__ == "__main__":
+    dir_path_in = '/home/alekseev_ilya/dialogue-augmentation/augmented/'
+    input_names = [
+        'back-translate',
+        'back-translate-cut',
+        'back-translate-shuffle',
+        'cut-insert',
+        'insert',
+        'pairwise-shuffler',
+        'pairwise-cutter',
+        'replace',
+        'replace-cut',
+        'shuffle-insert'
[…]

### 2023-10-22 src/mylib/dialogue_encoder/make_dataset.py

@@ -1,0 +1,123 @@
+original_path = '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/original'
+
+negative_paths = [
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/replace',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/replace-cut',
+]
+
+positive_paths = [
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/back-translate',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/back-translate-cut',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/back-translate-shuffle',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/cut-insert',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/insert',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/pairwise-shuffler',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/pairwise-cutter',
+    '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/augmented/shuffle-insert',
+]
+
+res_path = '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/dataset/train'
+
+N_CHUNKS = 80
+CHUNK_SIZE = 512
+ALLOW_ABSENT_NEGATIVES = True
+
+import json
+import os
+from tqdm import tqdm
+
+# === data validation ===
+
+def validate(paths):
+    all_chunk_names = []
+    for pos_path in tqdm(paths, desc='validating provided paths'):
+        chunk_names = [filename for filename in os.listdir(pos_path) if filename.endswith('.json') and not filename.startswith('ru')]
+        if len(chunk_names) != N_CHUNKS:
+            print('wrong number of chunks:', pos_path)
+        for chunk_name in chunk_names:
+            chunk_path = os.path.join(pos_path, chunk_name)
+            chunk = json.load(open(chunk_path, 'r'))
+            if len(chunk) != 512:
+                print('wrong chunk size:', chunk_name, pos_path)
+        all_chunk_names.append(chunk_names)
+
+    for chunk_names in all_chunk_names[1:]:
[…]

### 2023-10-22 src/mylib/dialogue_encoder/make_multiwoz_dataset.py

@@ -1,0 +1,53 @@
+CHUNK_SIZE = 512
+tokenizer = 'roberta-base'
+path_out = '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/multiwoz_truncated'
+
+if __name__ == "__main__":
+    from datasets import load_dataset
+    import json
+    import os
+    from filter_dataset_by_length import filter_dataset
+    from transformers import AutoTokenizer
+    from tqdm import tqdm
+    from filter_dataset_by_length import is_short_enough
+
+    dataset = load_dataset('multi_woz_v22')
+    tokenizer = AutoTokenizer.from_pretrained(tokenizer)
+    
+    def save_chunk(dataset_slice, path_out, i_chunk):
+        """Saves given slice as two json files: dia-{i_chunk}.json and services-{i_chunk}.json. Skips dias with exceeding length (see `is_short_enough()` in `filter_dataset_by_length.py`)"""
+        chunk_dias = []
+        chunk_services = []
+        for item in dataset_slice:
+            uts = item['turns']['utterance']
+            speakers = item['turns']['speaker']
+            dia = [{'utterance': ut, 'speaker': sp} for ut, sp in zip(uts, speakers)]
+
+            if not is_short_enough(dia, tokenizer):
+                print('rejected dia')
+                continue
+            
+            chunk_dias.append(dia)
+            services = item['services']
+            chunk_services.append(services)
+        
+        dia_path = os.path.join(path_out, f'dia-{i_chunk}.json')
+        services_path = os.path.join(path_out, f'services-{i_chunk}.json')
+        json.dump(chunk_dias, open(dia_path, 'w'))
+        json.dump(chunk_services, open(services_path, 'w'))
+
+    for split in ['train', 'validation', 'test']:
+        data_split = dataset[split]
+        
+        cur_path = os.path.join(path_out, split)
+        if not os.path.exists(cur_path):
+            os.makedirs(cur_path)
+        
+        split_size = len(data_split)
+        for i_chunk, start in tqdm(enumerate(range(0, split_size, CHUNK_SIZE)), desc=f'making chunks for {split}'):
+            end = start + CHUNK_SIZE
+            dataset_slice = [data_split[i] for i in range(start, min(end, split_size))]
+            
+            save_chunk(dataset_slice, cur_path, i_chunk)
+    
+    

### 2023-10-22 src/mylib/dialogue_encoder/train.py

@@ -1,0 +1,158 @@
+if __name__ == "__main__":
+    import torch
+    torch.set_float32_matmul_precision('medium')    
+
+    import argparse
+    ap = argparse.ArgumentParser()
+    ap.add_argument('--name', dest='name', default=None)
+    ap.add_argument('--cuda', dest='cuda', default='0')
+    ap.add_argument('--interval', dest='interval', required=True, type=int)
+    args = ap.parse_args()
+
+    # from dataclasses import dataclass
+    # @dataclass
+    # class Args:
+    #     name = None
+    #     cuda = '2'
+    #     interval = 100
+    # args = Args()
+
+    import os
+    os.environ['TOKENIZERS_PARALLELISM'] = 'false'
+    os.environ['CUDA_VISIBLE_DEVICES'] = args.cuda
+
+    import sys
+    sys.path.insert(0, '/home/alekseev_ilya/dialogue-augmentation/nup')
+    from models import SimpleDialogueEncoder
+    from train_utils import freeze_hf_model, Learner, LearnerConfig, ContrastiveDataset, MultiWOZServiceClfDataset
+
+    hf_model = 'roberta-base'
+    finetune_layers = 1
+
+    learner_config = LearnerConfig(
+        k=5,
+        t=0.1,
+        batch_size=32,
+        # warmup_period=200,
+        do_periodic_warmup=False,
+        lr=3e-6,
+        kwargs={
+            'finetune_layers': finetune_layers
+        }
+    )
+    
+    model = SimpleDialogueEncoder(hf_model)
+    freeze_hf_model(model.model, finetune_layers)
+
+    learner = Learner(model, learner_config)
+
+    # ======= DEFINE DATA =======
+
+    path = '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/dataset/train'
+    dataset = ContrastiveDataset(path)
+    
+    dir = '/home/alekseev_ilya/dialogue-augmentation/dialogue_encoder/multiwoz_truncated'
+    multiwoz_train = MultiWOZServiceClfDataset(
+        path=f'{dir}/train',
+        fraction=1.
+    )
+    multiwoz_val = MultiWOZServiceClfDataset(
+        path=f'{dir}/validation',
+        fraction=1.
+    )
+
+    from torch.utils.data import DataLoader
+    def collate_fn(batch):
+        return batch
+    train_loader = DataLoader(
+        dataset=dataset,
+        batch_size=learner_config.batch_size,
+        shuffle=False,
[…]

### 2023-10-22 src/mylib/dialogue_encoder/train_utils.py

@@ -1,0 +1,467 @@
+from torch.utils.data import Dataset
+import os
+import json
+import numpy as np
+from bisect import bisect_right
+from dataclasses import dataclass, field
+from typing import Tuple
+import lightning.pytorch as pl
+from torch.optim import AdamW
+from torch.optim.lr_scheduler import LambdaLR
+import torch.nn as nn
+import torch
+from typing import Literal
+import torch.nn.functional as F
+from torchmetrics.functional.classification import multilabel_f1_score
+from torch.utils.data import DataLoader
+import math
+from sklearn.neural_network import MLPClassifier
+from sklearn.metrics import f1_score
+
+
+class ContrastiveDataset(Dataset):
+    def __init__(self, path):
+        self.path = path
+        
+        chunk_names = [filename for filename in os.listdir(path) if filename.endswith('.json') and not filename.startswith('ru')]
+        self.chunk_names = sorted(chunk_names, key=lambda x: int(x.split('.')[0]))
+        chunk_sizes = [len(chunk) for chunk in (json.load(open(os.path.join(path, chunk_name))) for chunk_name in self.chunk_names)]
+        self.chunk_beginnings = np.cumsum(chunk_sizes).tolist()
+        
+        self.n_chunks = len(self.chunk_names)
+        self.len = self.chunk_beginnings[-1]
+    
+    def __len__(self):
+        return self.len
+    
+    def __getitem__(self, i):
+        """
+        Loads one chunk and returns one training sample as
+        {
+            'orig': dia,
+            'pos': list of dias,
+            'neg': list of dias
+        }
+        where each dia is represented with an object of the following schema:
+        ```
+        {
+            "type": "array",
+            "items":
+            {
+                "type": "object",
+                "properties":
+                {
+                    "utterance": {"type": "string"},
+                    "speaker": {"type": "number"}
+                }
+            }
+        }
+        ```"""
+        i_chunk = bisect_right(self.chunk_beginnings, x=i)
+        tmp = [0] + self.chunk_beginnings
+        idx_within_chunk = i -  tmp[i_chunk]
+        item = json.load(open(os.path.join(self.path, self.chunk_names[i_chunk]), 'r'))[idx_within_chunk]
+        return item
+
+
+class MultiWOZServiceClfDataset(Dataset):
+    def __init__(self, path, fraction=1.):
+        self.path = path
+        
[…]

### 2023-10-22 src/mylib/make_data_for_aug.py

@@ -1,0 +1,14 @@
+if __name__ == "__main__":
+    from datasets import load_dataset
+    import json
+
+    dataset = load_dataset("multi_woz_v22")
+    n_dialogues = 15
+    dialogues = []
+    for dia in dataset['validation']['turns'][:n_dialogues]:    
+        dialogues.append([{'utterance': ut, 'speaker': sp} for ut, sp in zip(dia['utterance'], dia['speaker'])])
+
+    import os
+    if not os.path.exists('aug-data'):
+        os.makedirs('aug-data')
+    json.dump(dialogues, open('aug-data/original.json', 'w'))

### 2023-10-22 src/mylib/my_augmentations.py

@@ -1,0 +1,1158 @@
+import pandas as pd
+import numpy as np
+import json
+import nltk
+from nltk.corpus import stopwords
+from transformers import AutoTokenizer, pipeline
+from typing import List, Literal
+from tqdm import tqdm
+
+def to_uts(dialogues):
+    res = []
+    for dia in dialogues:
+        for item in dia:
+            res.append(item['utterance'])
+    return res
+
+def to_dia(uts, dialogues):
+    dia_lens = [len(dia) for dia in dialogues]
+    res = []
+    for i, length in enumerate(dia_lens):
+        start = sum(dia_lens[:i])
+        end = start + length
+        cur_res = []
+        for j, ut in enumerate(uts[start:end]):
+            cur_res.append({
+                'speaker': dialogues[i][j]['speaker'],
+                'utterance': ut
+            })
+        res.append(cur_res)
+    return res
+
+
+class BackTranslator:
+    """Back Translate each utterance (separately). Preserves intent."""
+
+    def __init__(self, language, device='cpu'):
+        self.language = language
+        self.device = device
+
+    @staticmethod
+    def _load_utterances():
+        dialogues = json.load(open('aug-data/original.json', 'r'))
+        utterances = []
+        for dia in dialogues:
+            utterances.extend([item['utterance'] for item in dia])
+        return utterances, dialogues
+
+    @staticmethod
+    def _save(augmented, original, name):
+        """
+        Params
+        ------
+        - augmented: list of utterances
+        - original: list of lists of dicts with keys 'utterance', 'speaker'
+        - name: name of output .json file
+        """
+        i = 0
+        res = []
+        for dia in original:
+            aug_dia = []
+            for item in dia:
+                aug_dia.append({'utterance': augmented[i], 'speaker': item['speaker']})
+                i += 1
+            res.append(aug_dia)
+        json.dump(res, open(f'aug-data/{name}.json', 'w'))
+
+    def from_file_system(self, name='back_trans_hf'):
+        """
+        Params
+        ------
+        - name: str, name of output .json file
+        """
+
+        uts = self._load_utterances()
+        back_translated = self._augment(uts)
+        self._save(back_translated, dialogues, name)
+
+    def from_argument(self, dialogues):
+        """
+        Params
+        ------
+        - dialogues: list[list[str]]
+        """
+
+        uts = to_uts(dialogues)
+        forth = self._forth(uts)
+        back = self._back(forth)
+        
[…]

### 2023-10-22 src/mylib/nup/make_dataset.py

@@ -1,0 +1,130 @@
+import json
+from random import shuffle, seed
+from math import ceil
+import os
+from collections import defaultdict
+from tqdm import tqdm
+from datasets import load_dataset
+from transformers import AutoTokenizer
+
+
+names = [
+    'MS-DC',
+    'MetaLWOZ',
+    'MULTIWOZ2_2',
+    'SGD',
+    'SimJointGEN',
+    'KETOD',
+    'FRAMES',
+    'Disambiguation',
+    'ABCD',
+    'AirDialogue',
+    'BiTOD',
+    'Taskmaster1'
+]
+
+upper_bound = 96
+upper_bounds = {
+    'MS-DC': min(upper_bound, 250),
+    'MetaLWOZ': min(upper_bound, 100),
+    'MULTIWOZ2_2': min(upper_bound, 75),
+    # 'SGD': upper_bound,
+    'SimJointGEN': upper_bound,
+    'KETOD': upper_bound,
+    'FRAMES': upper_bound,
+    'Disambiguation': min(upper_bound, 60),
+    'ABCD': upper_bound,
+    'AirDialogue': upper_bound,
+    'BiTOD': upper_bound,
+    'Taskmaster1': min(upper_bound, 200),
+}
+
+tokenizer = AutoTokenizer.from_pretrained('microsoft/mpnet-base', max_length=10000)
+
+merged_dataset = []
+for name in names:
+    dataset = load_dataset('Salesforce/dialogstudio', name)['train']['log']
+    res = []
+    if name not in upper_bounds.keys():
+        for dia in tqdm(dataset, desc=name):
+            cur_dia = []
+            if len(dia) == 0 or len(dia) > 10 or len(dia) == 1 and (dia[0]['user utterance'] == '' or dia[0]['system response'] == ''):
+                continue
+            for turn in dia:
+                ut = turn['user utterance']
+                if ut != '':
+                    cur_dia.append({'utterance': ut, 'speaker': 0})
+                ut = turn['system response']
+                if ut != '':
+                    cur_dia.append({'utterance': ut, 'speaker': 1})
+            res.append(cur_dia)
+
+    else:
+        bound = upper_bounds[name]
+        for dia in tqdm(dataset, desc=name):
+            cur_dia = []
+            if len(dia) == 0 or len(dia) > 10 or len(dia) == 1 and (dia[0]['user utterance'] == '' or dia[0]['system response'] == ''):
+                continue
+            is_below_bound = []
+            for turn in dia:
[…]

### 2023-10-22 src/mylib/nup/models/__init__.py

@@ -1,0 +1,1 @@
+from .dialogue import SimpleDialogueEncoder

### 2023-10-22 src/mylib/nup/models/aux.py

@@ -1,0 +1,199 @@
+from torch import nn
+from transformers import AutoModel, AutoTokenizer
+from typing import List
+import torch
+import torch.nn.functional as F
+from dataclasses import dataclass
+import math
+
+
+class mySentenceTransformer(nn.Module):
+    """Imitation of SentenceTransformers (https://www.sbert.net/)"""
+
+    def __init__(
+            self,
+            model_name='sentence-transformers/all-mpnet-base-v2',
+            model=None,
+            pooling=True
+        ):
+        """If `pooling=False`, then instead of sentence embeddings forward will return list of token embeddings."""
+        super().__init__()
+        self.model_name = model_name
+        self.pooling = pooling
+
+        if model is None:
+            self.model = AutoModel.from_pretrained(model_name)
+        else:
+            self.model = model
+        self.tokenizer = AutoTokenizer.from_pretrained(model_name)
+
+    def forward(self, sentences: List[str]) -> List[torch.Tensor]:
+        input = self.tokenizer(sentences, padding='longest', return_tensors='pt').to(self.model.device)
+        output = self.model(
+            input_ids=input['input_ids'],
+            attention_mask=input['attention_mask']
+        )
+        
+        res = []
+        for token_emb, attention in zip(output.last_hidden_state, input['attention_mask']):
+            last_mask_id = len(attention)-1
+            while last_mask_id > 0 and attention[last_mask_id].item() == 0:
+                last_mask_id -= 1
+            embs = token_emb[:last_mask_id+1]
+            if self.pooling:
+                embs = torch.mean(embs, dim=0)
+                embs = embs / torch.linalg.norm(embs)
+            res.append(embs)
+
+        return res
+
+    def get_sentence_embedding_size(self):
+        return self.model.config.hidden_size
+
+
+class Projector(nn.Module):
+    """Fully-Connected 2-layer Linear Model. Taken from linking prediction paper code."""
+
+    def __init__(self, input_size, output_size):
+        super().__init__()
+        self.linear_1 = nn.Linear(input_size, input_size)
+        self.linear_2 = nn.Linear(input_size, input_size)
+        self.norm1 = nn.LayerNorm(input_size)
+        self.norm2 = nn.LayerNorm(input_size)
+        self.final = nn.Linear(input_size, output_size)
[…]

### 2023-10-22 src/mylib/nup/models/dialogue.py

@@ -1,0 +1,298 @@
+import torch
+from torch import nn
+from dataclasses import dataclass, asdict
+from .aux import myTransformerConfig, mySentenceTransformer, Projector, myTransformerBlock
+from .hssa import HSSAModel, HSSAConfig, HSSATokenizer
+from transformers import AutoModel, AutoTokenizer
+from transformers.models.mpnet.modeling_mpnet import create_position_ids_from_input_ids
+from collections import defaultdict
+from .train_utils import HParamsPuller
+from typing import Literal
+import torch.nn.functional as F
+
+
+@dataclass
+class UtteranceTransformerDMConfig(myTransformerConfig):
+    encoder_name: str = None
+    max_dia_len: int = None
+    embed_turn_ids: bool = True
+    is_casual: bool = False
+
+
+class UtteranceTransformerDM(nn.Module):
+    def __init__(self, config: UtteranceTransformerDMConfig):
+        super().__init__()
+
+        self.encoder = mySentenceTransformer(model_name=config.encoder_name)
+        self.max_dia_len = config.max_dia_len
+        self.embed_turn_ids = config.embed_turn_ids
+
+        sentence_embedding_dimension = self.encoder.get_sentence_embedding_size()
+        projection_size = sentence_embedding_dimension // 2
+        if self.embed_turn_ids:
+            self.hidden_size = projection_size + 16
+            self.turn_ids_embedding = nn.Embedding(config.max_dia_len, 8)
+        else:
+            self.hidden_size = projection_size + 8
+
+        self.speaker_embeddings = nn.Embedding(2, 8)
+        self.projector = Projector(sentence_embedding_dimension, projection_size)
+
+        config.hidden_size = self.hidden_size
+        config.intermediate_size = 4 * self.hidden_size
+        self.transformer = nn.ModuleList([myTransformerBlock(config) for _ in range(config.n_layers)])
+
+        self.config = config
+
+    def get_hparams(self):
+        return asdict(self.config)
+
+    @property
+    def device(self):
+        return self.encoder.model.device
+
+    def forward(self, batch):
+        dia_lens = [len(dia) for dia in batch]
+
+        inputs = []
+        for dia in batch:
+            speaker_ids = torch.tensor([item['speaker'] for item in dia], device=self.device, dtype=torch.long)
+            speaker_embeddings = self.speaker_embeddings(speaker_ids)
+
+            sentence_embeddings = self.encoder([item['utterance'] for item in dia])
[…]

### 2023-10-22 src/mylib/nup/models/hssa/__init__.py

@@ -1,0 +1,3 @@
+from .configuration_hssa import HSSAConfig
+from .modeling_hssa import HSSAModel
+from .tokenization_hssa import HSSATokenizer

### 2023-10-22 src/mylib/nup/models/hssa/modeling_hssa.py

@@ -1,0 +1,411 @@
+from transformers.models.mpnet.modeling_mpnet import (
+    MPNetSelfAttention as SelfAttention,
+    MPNetIntermediate as Intermediate,
+    MPNetOutput as Output,
+    MPNetPreTrainedModel as PreTrainedModel,
+)
+from typing import Optional
+from .configuration_hssa import HSSAConfig
+import torch.nn as nn
+import torch
+import math
+
+
+class SegmentPooler(nn.Module):
+    def __init__(self, config: HSSAConfig):
+        super().__init__()
+        self.hidden_size = config.hidden_size
+        self.dense = nn.Linear(self.hidden_size, self.hidden_size)
+        
+    def forward(self, token_states: torch.Tensor, attention_mask: torch.Tensor):
+        """
+        Pool segments
+
+        Params
+        ------
+            - `token_states`: (B*S, T, d)
+            - `attention_mask`: (B*S, T), attention mask for each utterance
+
+        Return
+        ------
+            - `utterance_states`: (B*S, d)
+            - `utterance_mask`: (B*S, 1)
+        
+        Notation
+        --------
+        B: batch size
+        S: segments number
+        d: hidden size
+        T: size of segment
+        """
+
+        _, T, hidden_size = token_states.shape
+
+        # (B*S, T, 1)
+        attention_mask = attention_mask.view(-1, T, 1)
+
+        # (B*S, d)
+        avg_tok_states = torch.sum(attention_mask * token_states, dim=1) / (1e-6 + attention_mask.sum(dim=1))
+
+        # (B*S, d, 1)
+        avg_tok_states = avg_tok_states.unsqueeze(2) 
+
+        # (B*S, T, d) x (B*S, d, 1) -> (B*S, T, 1)
+        scores = torch.bmm(token_states, avg_tok_states) / math.sqrt(hidden_size)
+        scores += attention_mask
+        scores = torch.softmax(scores, dim=1)
+
+        # (B*S, T, d) * (B*S, T, 1) -> (B*S, d)
+        utterance_states = torch.sum(token_states * scores, dim=1)
+        utterance_states = self.dense(utterance_states)
+
+        return utterance_states
+
+
+class SegmentUpdater(nn.Module):
+    def __init__(self, config: HSSAConfig):
+        super().__init__()
+        self.hidden_size = config.hidden_size
+
+    def forward(
+        self,
+        utterance_states: torch.Tensor,
+        token_states: torch.Tensor,
[…]

### 2023-10-22 src/mylib/nup/models/hssa/tokenization_hssa.py

@@ -1,0 +1,57 @@
+from transformers.models.mpnet.tokenization_mpnet_fast import MPNetTokenizerFast
+from transformers.tokenization_utils import AddedToken
+import torch
+
+
+class HSSATokenizer(MPNetTokenizerFast):
+    def __init__(
+        self,
+        sys_token='<system>',
+        user_token='<user>',
+        **kwargs
+    ):
+        # sending them as keyword argument `additional_special_tokens` will add these tokens to `additional_special_tokens`
+        additional_special_tokens = [
+            AddedToken(sys_token, lstrip=False, rstrip=False),
+            AddedToken(user_token, lstrip=False, rstrip=False),
+        ]
+
+        super().__init__(additional_special_tokens=additional_special_tokens, **kwargs)
+    
+        self.sys_token, self.user_token = self.additional_special_tokens
+        self.sys_token_id, self.user_token_id = self.additional_special_tokens_ids
+
+    # overrides Base's __call__
+    def __call__(self, text, **kwargs):
+        """
+        Input text is a list of dialogues, where each dialogue is a list of objects with following schema:
+        {
+            "type": "object",
+            "properties":
+            {
+                "utterance": {"type": "string"},
+                "speaker": {"type": "number"}
+            }
+        }
+        User and system speakers correspond to speaker==0 and speaker==1 respectively.
+        """
+        tokenizer = super().__call__
+        uts = []
+        uts_mask = []
+        max_dia_len = max(len(dia) for dia in text)
+
+        for dia in text:
+            for item in dia:
+                if item['speaker'] is not None:
+                    role_token = self.sys_token if item['speaker'] else self.user_token
+                else:
+                    role_token = ''
+                uts.append(role_token + item['utterance'])
+            for _ in range(max_dia_len - len(dia)):
+                uts.append('')
+            uts_mask.append([1] * len(dia) + [0] * (max_dia_len - len(dia)))
+        
+        tok_res = tokenizer(uts, padding='longest', return_tensors='pt')
+        tok_res['max_dia_len'] = torch.tensor([max_dia_len])
+        tok_res['utterance_mask'] = torch.tensor(uts_mask)
+        return tok_res

### 2023-10-22 src/mylib/nup/models/listwise.py

@@ -1,0 +1,457 @@
+import torch.nn as nn
+import torch
+import torch.nn.functional as F
+import numpy as np
+from bisect import bisect_left
+from .train_utils import LightningCkptLoadable, HParamsPuller
+from typing import Literal
+import warnings
+
+
+# ======= ranker =======
+
+class RankerHead(nn.Module, HParamsPuller):
+    def __init__(self, hidden_size, dropout_prob):
+        super().__init__()
+
+        self.lin = nn.Linear(hidden_size, hidden_size)
+        self.dropout = nn.Dropout(dropout_prob)
+        self.ranker = nn.Linear(hidden_size, 1, bias=False)
+    
+    def forward(self, x: torch.Tensor):
+        x = F.gelu(self.lin(x)) + x
+        x = self.dropout(x)
+        x = self.ranker(x)
+        return x.squeeze(-1)
+
+
+class SortingLoss(nn.Module):
+    def __init__(self):
+        super().__init__()
+        self.loss_fn = nn.KLDivLoss(reduction='batchmean')
+
+    def forward(self, ranks_logits, dia_lens):
+        ranks_logprobs = F.log_softmax(ranks_logits, dim=1)
+        _, T = ranks_logits.shape
+        device = ranks_logits.device
+        ranks_true = self._make_true_ranks(T, dia_lens, device)
+        return self.loss_fn(ranks_logprobs, ranks_true)
+
+    @staticmethod
+    def _make_true_ranks(T, dia_lens, device):
+        res = []
+        
+        for length in dia_lens:
+            ranks = torch.linspace(1, 0, length, device=device)
+            ranks = F.pad(ranks, pad=(0, T-length), value=0)
+            ranks = ranks / ranks.sum()
+            res.append(ranks)
+        
+        return torch.stack(res)
+
+
+class SortingMetric:
+    def __call__(self, ranks_logits, mask, dia_lens):
+        unbinded_ranks_logits = self._unbind_logits(ranks_logits, mask, dia_lens)
+        permutations = self._to_permutations(unbinded_ranks_logits)
+        return 1-np.mean([self._normalized_inversions_count(perm) for perm in permutations])
+       
+    @staticmethod
+    def _unbind_logits(logits, mask, dia_lens):
+        """get list of tensors with logits corresponding to tokens(utterances) that are not padding ones only"""
+        return logits[~mask].detach().cpu().split(dia_lens)
+
+    @staticmethod
+    def _to_permutations(unbinded_ranks_logits):
[…]

### 2023-10-22 src/mylib/nup/models/pairwise.py

@@ -1,0 +1,260 @@
+import torch.nn as nn
+from .train_utils import LightningCkptLoadable, HParamsPuller
+import torch
+import torch.nn.functional as F
+import numpy as np
+from .aux import mySentenceTransformer, Projector
+
+
+class ChainCosine(nn.Module, LightningCkptLoadable, HParamsPuller):
+    def __init__(self, target_encoder: mySentenceTransformer, context_encoder, projection_size, context_size, temperature=1, k=1, hard_negative=False):
+        super().__init__()
+
+        self.projection_size = projection_size
+        self.context_size = context_size
+        self.temperature = temperature
+        self.k = k
+        self.hard_negative = hard_negative
+
+        self.target_encoder = target_encoder
+        self.context_encoder = context_encoder
+        
+        self.context_projector = Projector(
+            input_size=self.context_encoder.get_encoding_size(),
+            output_size=self.projection_size
+        )
+        self.target_projector = Projector(
+            input_size=self.target_encoder.get_encoding_size(),
+            output_size=self.projection_size
+        )
+
+    @property
+    def device(self):
+        return self.target_encoder.model.device
+    
+    def get_encodings(self, batch):
+        if self.context_size is None:
+            context_slice = slice(None, None, None)
+        else:
+            context_slice = slice(-self.context_size, None, None)
+
+        context_batch = []
+        target_batch = []
+        for pair in batch:
+            context_batch.append(pair['context'][context_slice])
+            target_batch.append(pair['target'])
+        
+        target_encodings = self.target_encoder(target_batch)
+        context_encodings = self.context_encoder(context_batch)
+
+        context_encodings = self.context_projector(context_encodings)
+        target_encodings = self.target_projector(target_encodings)
+
+        return context_encodings, target_encodings
+    
+    def get_logits(self, batch, temperature):
+        context_encodings, target_encodings = self.get_encodings(batch)
+        return context_encodings @ target_encodings.T / temperature
+    
+    def forward(self, batch):
+        logits = self.get_logits(batch, self.temperature)
+        
+        if self.hard_negative:
+            mask = torch.eye(len(batch), dtype=torch.bool, device=logits.device)
+            
+            positive = torch.exp(logits[mask])
[…]

### 2023-10-22 src/mylib/nup/models/train_utils.py

@@ -1,0 +1,275 @@
+from torch.utils.data import Dataset
+from typing import Literal, Tuple
+import math
+import json
+from dataclasses import dataclass, field
+import lightning.pytorch as pl
+import torch.nn as nn
+from torch.optim import AdamW
+from torch.optim.lr_scheduler import LambdaLR
+
+
+class NUPDataset(Dataset):
+    chunk_size = 2048
+    def __init__(self, path, split: Literal['train', 'test', 'val'], fraction=1.):
+        self.split = split
+        self.path = path
+
+        if split == 'train':
+            max_n_chunks = 2556
+        elif split == 'test' or split == 'val':
+            max_n_chunks = 141
+
+        if isinstance(fraction, float):
+            self.fraction = min(1., max(0., fraction))
+            self.n_chunks = math.ceil(self.fraction * max_n_chunks)
+        elif isinstance(fraction, int):
+            self.fraction = min(max_n_chunks, max(1, fraction))
+            self.n_chunks = fraction
+        else:
+            raise ValueError('fraction must be int or float')
+
+        self.len = self.n_chunks * self.chunk_size
+    
+    def __len__(self):
+        return self.len
+    
+    def __getitem__(self, i):
+        """
+        Loads one chunk and returns one dialogue, represented with an object of the following schema:
+        ```
+        {
+            "type": "object",
+            "properties":
+            {
+                "context":
+                {
+                    "type": "array",
+                    "items":
+                    {
+                        "type": "object",
+                        "properties":
+                        {
+                            "utterance": {"type": "string"},
+                            "speaker": {"type": "number"}
+                        }
+                    }
+                },
+                "target":
+                {
+                    "type": "object",
+                    "properties":
+                    {
+                        "utterance": {"type": "string"},
+                        "speaker": {"type": "number"}
+                    }
+                }
+            }
+        }
+        ```"""
+        i_chunk = math.floor(i / self.chunk_size)
+        idx_within_chunk = i % self.chunk_size
+        item = json.load(open(f'{self.path}/pairs/{self.split}/{i_chunk}.json', 'r'))[idx_within_chunk]
+        return item
+
+
+class DialogueDataset(Dataset):
+    chunk_size = 512
+    def __init__(self, path, split: Literal['train', 'test', 'val'], fraction=1.):
+        self.split = split
+        self.path = path
+
[…]

### 2023-10-22 src/mylib/nup/train.py

@@ -1,0 +1,466 @@
+if __name__ == "__main__":
+    
+    # ======= DEFINE TASK =======
+
+    import torch
+    torch.set_float32_matmul_precision('medium')    
+
+    import argparse
+    ap = argparse.ArgumentParser()
+    ap.add_argument('--model', dest='model', required=True, choices=[
+        'pairwise-cat', 'pairwise-ema', 'pairwise-sparse-transformer',
+        'listwise-utterance-transformer', 'listwise-sparse-transformer',
+        'listwise-hssa', 'listwise-clf', 'listwise-decoder', 'pairwise-cat-decoupled',
+        'pairwise-ema-both'
+    ])
+    ap.add_argument('--name', dest='name', default=None)
+    ap.add_argument('--cuda', dest='cuda', default='0')
+    ap.add_argument('--interval', dest='interval', required=True, type=int)
+    args = ap.parse_args()
+
+    import os
+    os.environ['TOKENIZERS_PARALLELISM'] = 'false'
+    os.environ['CUDA_VISIBLE_DEVICES'] = args.cuda
+
+    # ======= DEFINE LEARNER =======
+
+    from models.hssa.modeling_hssa import SegmentPooler
+    def freeze_hssa(model, finetune_layers=0):
+        model.embeddings.requires_grad_(False)
+        model.embeddings.word_embeddings.weight[-2:].requires_grad_(True)
+
+        model.encoder.requires_grad_(False)
+        for i, layer in enumerate(model.encoder.layer):
+            layer.requires_grad_(i>=model.config.num_hidden_layers-finetune_layers)
+
+        for module in model.modules():
+            if isinstance(module, SegmentPooler):
+                module.requires_grad_(True)
+
+    def freeze_hf_model(hf_model, finetune_layers):
+        """Freeze all encoder layers except last `finetune_encoder_layers`"""
+        hf_model.requires_grad_(False)
+        n_layers = hf_model.config.num_hidden_layers
+        for i in range(n_layers):
+            hf_model.encoder.layer[i].requires_grad_(i>=n_layers-finetune_layers)
+
+    mpnet_name = 'sentence-transformers/all-mpnet-base-v2'
+    amazon_name = 'aws-ai/dse-roberta-base'
+
+    from models.train_utils import Learner, LearnerConfig
+
+    from models.aux import mySentenceTransformer
[…]

### 2023-10-22 src/mylib/preparation_for_similarity_analisys.sh

@@ -1,0 +1,6 @@
+source VENV/bin/activate
+export OS_CUDA_VISIBLE_DEVICES="0"
+python3 parse_multiwoz.py
+python3 sentence_encoding.py
+python3 dgac_clustering.py
+python3 assign_clust_tf_idf_names.py

### 2023-10-22 src/mylib/sentence_encoding.py

@@ -1,0 +1,23 @@
+from sentence_transformers import SentenceTransformer
+import numpy as np
+
+
+def sentence_encoder(utterances):
+    """
+    Encodes `utterances` with sentence_transformers library and saves embeddings to .npy file.
+
+    Params
+    ------
+        utterances: list[str], all utterances from dataset
+        path: str, where to save .npy file
+    """
+    model = SentenceTransformer('sentence-transformers/all-distilroberta-v1').to('cuda')
+    return model.encode(utterances, show_progress_bar=True)
+
+
+if __name__ == '__main__':
+    import json
+
+    utterances = json.load(open('clust-data/utterances.json', 'r'))
+    
+    np.save('clust-data/sentence_embeddings', sentence_encoder(utterances))

### 2023-10-22 src/mylib/similarity_functions.py

@@ -1,0 +1,98 @@
+import numpy as np
+import json
+from sentence_encoding import sentence_encoder
+from dgac_clustering import Clusters
+import pickle
+
+
+def count_vectorizer_from_sile_system(names):
+    """
+    1. Encodes each utterance with sentence encoder specified in sentence_encoding.py
+    2. Infers DGAC clusterer to predict labels for each utterance
+    3. Saves them to json as f'aug-data/clust-labels-{name}.json'
+    4. Vectorizes each dialogue as bag of cluster labels (analogue to bag of words)
+    5. Saves vectorizations to .npy files as f'aug-data/vectors-{name}'
+
+    Params
+    ------
+    - names: list[str], is used to read original and augmented dialogues in format f'aug-data/{name}.csv'
+    """
+    clusterer: Clusters = pickle.load(open('clust-data/dgac_clusterer.pickle', 'rb'))
+
+    for name in names:
+        dialogues = json.load(open(f'aug-data/{name}.json', 'r'))
+        utterances = []
+        speaker = []
+        rle = []
+        for dia in dialogues:
+            i = len(utterances)
+            for item in dia:
+                utterances.append(item['utterance'])
+                speaker.append(item['speaker'])
+            rle.append(len(utterances) - i)
+        speaker = np.array(speaker)
+        embeddings = sentence_encoder(utterances)
+        
+        labels_user, labels_system = clusterer.predict(embeddings, speaker)
+        labels = np.empty(len(embeddings), dtype=np.int16)
+        labels[speaker == 0] = labels_user
+        labels[speaker == 1] = labels_system + clusterer.n_clusters // 2
+
+        vectors = []
+        labels_list = []
+        for i in range(len(rle)):
+            start = sum(rle[:i])
+            end = start + rle[i]
+            vectors.append(np.bincount(labels[start:end], minlength=clusterer.n_clusters))
+            labels_list.append(labels[start:end].tolist())
+        
+        np.save(f'aug-data/vectors-{name}', vectors)
+        json.dump(labels_list, open(f'aug-data/clust-labels-{name}.json', 'w'))
+
+
+def count_vectorizer_from_argument(augmented_utterances, rle, speaker, clusterer):
+    """
+    1. Encodes each utterance with sentence encoder specified in sentence_encoding.py
[…]

### 2023-10-22 src/mylib/visualization_utils.py

@@ -1,0 +1,168 @@
+import plotly.express as px
+from sklearn.manifold import TSNE
+import pickle
+from dgac_clustering import Clusters
+import json
+import numpy as np
+import pandas as pd
+from redlines import Redlines
+from IPython.display import Markdown, display
+from Levenshtein import distance, ratio, jaro
+from typing import Literal
+from difflib import HtmlDiff
+import html2markdown
+import os
+
+
+def show_clusters(is_system) -> None:
+    """
+    Show clusters, obtained from train multiwoz with DGAC clustering. Scatter plot with utterances appearing on hover.
+    Also assigns names to clusters using TF-IDF and saves them to clust-data directory as .json files.
+
+    Params
+    ------
+        is_system: bool, flag, indicating what part of clusters to visualize, user of system
+    """
+
+    # load necessary data
+    clusterer: Clusters = pickle.load(open('clust-data/dgac_clusterer.pickle', 'rb'))
+    X = np.load('clust-data/sentence_embeddings.npy')
+    speaker = np.array(json.load(open('clust-data/speaker.json', 'r')))
+
+    mask = (speaker == is_system)
+        
+    projected = pd.DataFrame(TSNE().fit_transform(X[mask]))
+
+    # get names of clusters via tf-idf
+    texts = [txt for i, txt in enumerate(json.load(open('clust-data/utterances.json', 'r'))) if speaker[i] == is_system]
+    labels = clusterer.labels[speaker == 1] - is_system * clusterer.n_clusters // 2
+    apndx = 'system' if is_system else 'user'
+    cluster_names = json.load(open(f'clust-data/cluster-tfidf-names-{apndx}.json', 'w'))
+
+    projected['clust_name'] = '-'
+    for i in range(clusterer.n_clusters // 2):
+        projected.loc[labels == i, 'clust_name'] = cluster_names[i]
+    
+    # visualize
+    projected['label'] = labels
+    projected['text'] = texts
+    fig = px.scatter(
+        projected, x=0, y=1, color='label',
+        hover_name='clust_name',
+        hover_data={'label': True, 'text': True},
+        title=f'{apndx} utterances'
+    )
+
+    fig.update_layout()
+    fig.show()
+
+
+def show_augmented(i, name) -> None:
+    """
+    Show difference between original dialogue and augmented.
+    
+    Params
+    ------
[…]

### 2023-11-08 paper/main.tex
added related works

@@
 \usepackage{hyperref}       % hyperlinks
+\usepackage[dvipsnames]{xcolor}
 \usepackage{url}            % simple URL typesetting
@@
 \graphicspath{{figures/}}     % organize your images and other figures under figures/ folder
+
+\hypersetup{
+    colorlinks=true,
+    linkcolor=teal,
+    % filecolor=magenta, 
+    urlcolor=magenta,
+    citecolor=teal,
+    % pdftitle={GD},
+    % pdfpagemode=FullScreen,
+    % linktoc=all
+    }
 
@@
 %% Title
-\title{Generating Synthetic Data via Intent-Preserving and Intent-Corrupting Augmentations for Multi-task Training Dialogue Embeddings.
+\title{Contrastive Learning with Intent-Preserving and Intent-Corrupting Augmentations for Training Dialogue Embeddings.
 %%%% Cite as
@@
   Ilya Alekseev \\
-  Neural Networks and Deep Learning Lab \\
-  Moscow Institute of Physics and Technology \\
+  Lomonosov Moscow State University \\
   \texttt{ilya\_alekseev\_2016@list.ru} \\
@@
   Denis Kuznetsov \\
-  Neural Networks and Deep Learning Lab \\
   Moscow Institute of Physics and Technology \\
@@
 Unfortunately, there are no such datasets for the domain of dialogue data.
-In this work, we describe the process of mining a synthetic dataset of dialogues for contrastive learning with hard negatives.
-We investigate various augmentation strategies for constructing dialogues with preserved or corrupted intents (positive and negative samples, respectively).
-To demonstrate the stated cleanliness and diversity, we train a dialogue encoder model and analyze its properties.
+In this work, we describe the process of mining a synthetic dataset of dialogues for contrastive learning with hard negative samples.
+We apply various augmentation strategies for constructing dialogues with preserved or corrupted sets of intents.
+We train dialogue embeddings and report its performance on transfer learning tasks: domain classification, intent similarity, dialogue retrieval.
 \end{abstract}
@@
 % keywords can be removed
-\keywords{text embedding \and dialogue \and synthetic data \and augmentation \and natural language processing}
+\keywords{text embedding \and dialogue \and augmentation \and natural language processing \and contrastive learning}
 
 \section{Introduction}
-Obtaining embeddings is one of the key tasks in machine learning and popular one in recent years.
-Vector representation of an object is a convenient mathematical object.
-If an embedding accurately and comprehensively encodes the semantics of the original data, it opens up the possibility of using it in a wide range of tasks.
+Obtaining embeddings is one of the key tasks in machine learning and popular one in recent years, because vector representation of an object is a convenient mathematical object.
+If an embedding accurately and comprehensively encodes the semantics of the original data, it opens up the possibility of using it in a wide range of downstream tasks.
 
 In the field of natural language processing, classical methods for text vectorization such as bag of words \cite{bow} and tf-idf \cite{SprckJones2021ASI} have long been discovered.
-Thanks to deep learning, we have witnessed remarkable word vectorizations such as word2vec \cite{mikolov2013efficient} and GloVe \cite{pennington-etal-2014-glove}, which convey the semantic similarity between words; CoVe \cite{mccann2018learned} and ELMo \cite{peters-etal-2018-deep}, which encode information about the surrounding context.
+Thanks to deep learning, we have witnessed remarkable word vectorization techniques such as word2vec \cite{mikolov2013efficient} and GloVe \cite{pennington-etal-2014-glove}, which convey the semantic similarity between words; CoVe \cite{mccann2018learned} and ELMo \cite{peters-etal-2018-deep}, which encode information about the surrounding context.
 Recently, powerful encoder models have emerged that produce general-purpose text embeddings \cite{xiao2023cpack, wang2022text, li2023general}.
-They incorporate so much semantic information about texts that it can be applied to tasks such as classification, clustering, ranking, semantic textual similarity, and more.
+They incorporate so much semantic information about texts that it can be applied to tasks such as classification, clustering, ranking, semantic textual similarity, and more \cite{muennighoff2023mteb}.
 The success of these models is largely attributed to the use of contrastive learning on massive datasets.
@@
 However, the more specific the data structure, the more challenging it is to mine a large dataset.
-As of today, there are no encoding methods for entire dialogues.
-In other words, there is no way to obtain a dense vector representation that conveys universal information about a dialogue.
-There are language models that adapt successfully to the hierarchical and temporal nature of dialogue \cite{zhang-etal-2023-dialog, li2022future}, but they only address tasks at the token and utterance levels, not at the dialogue level.
-There are text encoders for utterances in a dialogue \cite{henderson2020convert, Zhou2022}, but not for the entire dialogue.
+Language models, such as those presented in \cite{zhang-etal-2023-dialog, li2022future}, have demonstrated successful adaptation to the hierarchical and temporal characteristics of dialogues.
+However, it is important to note that their training primarily involves inter-token and inter-utterance tasks rather than inter-dialogue tasks.
 
 Data is almost always scarce when it comes to building dialogue models.
-To date, numerous methods for generating synthetic dialogues have been devised \cite{kim2021neuralwoz, mohapatra2021simulated, wan-etal-2022-unified, zheng2023augesc}, but they do not generate dialogues in pairs \cite{schick2021generating}, as is conceptually important for contrastive learning.
-The simplest way to expand a training dataset is through augmentation \cite{soudani2023data}.
-However, we find them insufficient, as they do not significantly alter the structure of the dialogue.
-In this work, we will describe a method for constructing a synthetic dataset of dialogue pairs using various augmentations that preserve or alter the set of intents in the dialogue.
-These augmentations can be used for contrastive learning together with other tasks for pretraining powerful dialogue embeddings.
+To date, numerous methods for generating synthetic dialogues have been devised \cite{kim2021neuralwoz, mohapatra2021simulated, wan-etal-2022-unified, zheng2023augesc, schick2021generating}, but they do not generate dialogues in pairs, as it is conceptually important for training SoTA text embedding.
+The simplest way to generate pairs is through augmentation \cite{soudani2023data}.
+In this work, we will describe a method for constructing a dialogue dataset for contrastive learning using various augmentations that preserve or alter the set of intents in the dialogue.
+These augmentations can be used for contrastive learning for pretraining powerful dialogue embeddings.
 
@@
 where $u_i$ represents the utterance of a participant in the dialogue at step $s_i$.
-We are interested in so-called task-specific dialogues with two participants: the system and the user.
+We are interested in so-called task-oriented dialogues with two participants: the system and the user.
 With some approximation, they can be described as dialogues between a customer and a service worker (or a robot).
@@
 Let $D$ be the set of valid objects.
-Then augmentation is a non-identical mapping $\text{aug}(d)$ that does not take objects outside the set of valid objects:
-$$
-\text{aug}: D\to D.
-$$
+Then augmentation is a non-identical mapping $\text{aug}(d)$ that does not take objects outside the set of valid objects: $\text{aug}: D\to D.$
+
 Typically, this generation is implemented by making small changes to a valid training object.
@@
 Text augmentation is especially challenging because validity implies adherence to language rules, meaningfulness and a certain style.
-In the case of dialogues, it is necessary to maintain the structure and role differentiation, as indicated earlier.
+In the case of dialogues, it is necessary to maintain the structure and role differentiation.
 
@@
 
-\texttt{this section is incomple, this is still a draft }
+\texttt{this section is incomplete, this is still a draft }
 
-AugSBERT \cite{thakur2021augmented} can be viewed as the similar approach to ours one, but their augmentation uses an already presented dataset of text pairs, which is absent in our case.
+\subsection{Text Embedding}
 
-\section{Approach}
+One of the breakthroughs in text embedding came from the need for solving the task of semantic textual similarity \cite{reimers2019sentencebert}, i.e.
+comparing texts with each other.
+Previously existing methods were either weak (averaging word embeddings) or required enormous amount of computation (BERT cross-encoder).
+Comparing texts with bi-encoder is performed via averaging hidden states from the last transformer block of language model or taking hidden state of special token \texttt{[CLS]}.
+This setting opens up a possibility for clustering, retrieval, pair classification etc.
+within text domain \cite{muennighoff2023mteb}.
+
+\subsection{Contrastive Learning}
+
+To make the encoder produce rich vector representations, it is necessary to train it with tasks where semantic features are engaged. A popular one is contrastive learning:
+$$
+\mathcal{L}=-\log{\exp(\text{sim}(x,y))\over\sum_{z}\exp(\text{sim}(x,z))}.
+$$
+Here, $x=e_\theta(a), y=e_\theta(b)$ are embeddings of a pair of semantically similar objects and $z=e_\theta(c)$ is semantically distant from $x$, sim is a metric similarity function (e.g.
+cosine).
+This task trains model parameters $\theta$ to make embeddings convey semantic similarity as cosine similarity.
+
+This objective can be viewed as a special case of self-supervised learning, that long been discovered in the field of computer vision \cite{noroozi2017unsupervised, doersch2017multitask, wu2018unsupervised, oord2019representation}.
+This unsupervised approaches introduced a new view of transfer learning, allowing to advance from learning models to learning representations, which are more convenient to use for downstream tasks.
+
+% The main problem for contrastive learning is how to mine positive pairs and negative samples.
+One optimization step of contrastive learning can be viewed as a motion in the embedding space, in which positive objects attract to and negative pairs move away from each other.
+If positive and negative samples are not representative, this motion is senseless or inefficient.
+
+One of the most well-known methods for negative sampling was introduced by word2vec \cite{mikolov2013efficient} and it is to make random sampling from the dataset.
+Present-day methods with random negative sampling (e.g.
+OpenAI embeddings \cite{neelakantan2022text}) implement this idea as in-batch negative sampling and rely on large batch.
+Another idea is to use hard negative samples \cite{xiong2020approximate, feng2022languageagnostic}, which are the closest to positive pair sample among all negative samples.
+
+% \begin{figure}[!ht]
+%     \begin{subfigure}[t]{0.25\linewidth}
+%         \includegraphics[width=1\linewidth]{figures/in-batch.drawio.pdf}
+%         \caption{in-batch}
+%         \label{fig:in-batch}
+%     \end{subfigure}
+%     \begin{subfigure}[t]{0.25\linewidth}
+%         \includegraphics[width=1\linewidth]{figures/large-batch.drawio.pdf}
+%         \caption{large batch}
+%         \label{fig:large-batch}
+%     \end{subfigure}
+%     \begin{subfigure}[t]{0.25\linewidth}
+%         \includegraphics[width=1\linewidth]{figures/hard-negatives.drawio.pdf}
+%         \caption{hard negatives}
+%         \label{fig:hard-negatives}
+%     \end{subfigure}
+%     \centering
+%     \caption{Negative sampling. Conceptual view on contrastive learning.}
+% \end{figure}
+
+The most efficient way of mining positive pairs is to use supervision.
+For example, datasets on NLI (natural language inference), QA (question answering), fact verification, paraphrases contain pairs of semantically similar texts and are appropriate for training text embedding \cite{li2023general}.
+This data is extremely clean, but scarce.
+Another way is to scrap data from web pages, such as QA forums (Quora), social media (Reddit), etc.
+This data is large, but can be noisy.
+When there are no any facility like supervised or scrapped pairs, augmentation is the last variant.
+Two augmented views of the same object must preserve semantics, that are essential for this object, but provide sufficiently hard task for a model of discriminating one positive sample from the rest of negative ones.
+
+Present-day SoTA text embeddings such as BGE \cite{xiao2023cpack}, GTE \cite{li2023general}, E5 \cite{wang2022text} follow one pipeline of training.
+First, a retrieval-oriented language model is trained \cite{xiao2022retromae, gao-callan-2021-condenser}, that is a BERT-like model that can efficiently aggregate global information about input sequence.
+Second, a general-purpose fine-tuning is performed on a large weakly supervised text pairs with large batch contrastive learning.
+Third, task-specific fine-tuning is performed on supervised text pairs with
+contrastive learning with hard negative samples.
+
+\subsection{Similar Approaches}
+
+AugSBERT \cite{thakur2021augmented} cannot be viewed as the similar approach to ours. Its augmentation uses an already presented dataset of text pairs, which is absent in our case.
+
+\textbf{Text embedding.} CERT \cite{fang2020cert} utilizes back-translation for mining positive pairs.
+ConSERT \cite{yan-etal-2021-consert} utilizes various augmentations at the token-level.
+Doc2vecC \cite{luo2021unsupervised} utilizes thesaurus-based augmentations and back-translation.
+
+\textbf{Dialogue embedding.} Dial2vec \cite{liu2022dial2vec} modifies transformer architecture by adding cross-attention between utterances of different participants.
+DialogueCSE \cite{liu2021dialoguecse} adds cross-attention between utterances of different stages of a dialogue.
+
+Our approach is distinct from these above:
+\begin{itemize}
+    \item our augmentations do not take objects outside the set of valid objects (do not violate grammars and meaningfulness), because our augmentations are thoroughly designed with context-aware language models in contrast to ConSERT that simply shuffles and deletes tokens leaving only lexical info in touch
+    \item our augmentations form a full-fledged set of transformations that construct non-obvious examples, not just paraphrasing or token-level replacements
+    \item our approach doesn't modify BERT architecture
+\end{itemize}
+
+\section{Method}
+
+In this section, we describe our method.
+It can be divided into two stages.
+At the first stage, we develop and make augmentations for dialogue data.
+In result, the dataset consists of samples with following schema: original dialogue, set of augmentations with preserved intents, set with corrupted intents.
+At the second stage, we apply the contrastive learning.
 
@@
 \textbf{Token Insertion.} One of the simple yet effective ways to augment text is to lengthen it by inserting extra tokens.
-For this purpose, we added a special token `<mask>` to random places in the dialogues and used transformer models trained on the MLM task to fill these masks \cite{devlin2019bert}.
+For this purpose, we added a special token \texttt{<mask>} to random places in the dialogues and used transformer models trained on the MLM task \cite{devlin2019bert} to fill these masks.
 Insertion is rejected if the token proposed by the model is only a fragment of a word \cite{wu2016googles, sennrich-etal-2016-neural} or if the prediction probability is below a manually set threshold.
@@
 
-\textbf{Token Replacement.} This method is identical to the previous one, except that instead of adding the `<mask>` token, some tokens in the original dialogue are replaced.
+\textbf{Token Replacement.} This method is identical to the previous one, except that instead of adding the \texttt{<mask>} token, some tokens in the original dialogue are replaced.
 In this case, the mask-filling model is fed with single utterances to make replacements more diverse and random.
@@
 
-To expand the set of augmentations even further, we use a composition of several. More details about augmentation implementations are presented in appendix.
+To expand the set of augmentations even further, we use a composition of several ones. For more details on augmentation implementations, please refer to Appendix \ref{app:aug}.
 
@@
 
-As a baseline method for vectorization, we use RoBERTa \cite{liu2019roberta} without any modifications.
-The input is \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3 [SEP]}, and the output is the hidden state of \texttt{[CLS]} token from the last layer.
+As a model for embedding, we try BERT \cite{devlin2019bert}, RoBERTa \cite{liu2019roberta}, RetroMAE \cite{xiao2022retromae} without any modifications.
+The input is \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3}, and the output is the hidden state of \texttt{[CLS]} token from the last layer.
 
-For a slightly advanced dialogue language model, we use HSSA \cite{zhang-etal-2023-dialog}.
-It uses BERT \cite{devlin2019bert} as a backbone and modifies its attention to capture the hierarchical structure of dialogue and reach the computational trade-off between feeding a transformer separate utterances and feeding an entire dialogue.
+Also, we experiment with a slightly advanced dialogue language model, HSSA \cite{zhang-etal-2023-dialog}.
+It uses BERT as a backbone and modifies its attention mechanism to capture the hierarchical structure of dialogue and reach the computational trade-off between feeding a transformer separate utterances and feeding an entire dialogue.
 
-\subsection{Pre-train Tasks}
+\subsection{Pre-train Task} \label{sec:pretrain}
 
-To make the encoder produce rich vector representations, it is necessary to train it to perform tasks where semantic features are engaged. A popular task is contrastive learning:
-$$
-\mathcal{L}=-\log{\exp(\text{sim}(x,y))\over\sum_{z}\exp(\text{sim}(x,z))}.
-$$
-Here, $x=e_\theta(a), y=e_\theta(b)$ are embeddings of a pair of semantically similar objects and $z=e_\theta(c)$ is semantically distant from $x$, $\text{sim}$ is a metric similarity function.
-This task trains embeddings to convey semantic similarity as metric similarity.
+In our case, $a, b$ are two intent-preserving augmentations of the same dialogue, and the remaining dialogues from the training batch are used as negative examples.
+This task simulates the second stage of training general-purpose text embeddings, which uses a large batch and weakly supervised text pairs.
+In addition to this, in order to reduce required batch size, we utilize intent-corrupting augmentations as hard negative samples.
+Notably, the technique with two augmentations and dual encoder resembles BYOL \cite{grill2020bootstrap}, well-known method for self-supervised learning \texttt{(maybe move this sentence to Related Works)}.
+
+\subsection{Evaluation}
+
+We evaluate dialogue embeddings in transfer learning setting.
+Specifically, our evaluation methods utilize embeddings as features, without modifying the encoder model itself.
+Inspired by \cite{liu2022dial2vec} we employ these evaluation methods: 1) domain classification, 2) dialogue retrieval, 3) intent similarity.
+Evaluation is performed on MultiWOZ 2.2 dataset \cite{zang2020multiwoz}.
+
+\textbf{Domain classification.} The goal is to predict in which domain a dialogue is taking place.
+In dataset there are 7 domains: attraction, bus, hospital, hotel, restaurant, taxi, train.
+Each dialogue can take place in multiple domains at once.
+The method is to train a linear classifier upon dialogue embedding, this is the so-called linear probe evaluation, that is used in many works.
+F1-macro is used to measure quality.
+
+\textbf{Dialogue retrieval.} For each dialogue from validation split, the goal is to retrieve dialogues from train split with at least one domain in common.
+Ranking score is calculated as cosine similarity between query and answer embeddings.
+Mean average precision at 100 is used to measure quality.
+
+\textbf{Intent similarity.} We sample pairs of dialogues from train and validation splits of dataset.
+A linear classifier is trained on the former pairs and evaluated on the latter pairs using Pearson correlation between predicted similarity scores and gold ones.
+The gold scores are obtained using DGAC clustering \cite{dgac}.
+Clusters represent intents of dialogue participants.
+Hence, each dialogue can be associated with a set of intents.
+We define gold intent similarity of two dialogues as a dice similarity between their sets of intents.
+
+\section{Experiments}
+
+\textbf{Dataset.} In all experiments, we used the same dataset of dialogues.
+This dataset is a combination of several task-oriented dialogue datasets from DialogStudio \cite{zhang2023dialogstudio}.
+All the dialogues were filtered based on their length, resulting in a dataset comprising 501K dialogues.
+For more details on the dataset used, please refer to Appendix \ref{app:dataset}.
+
+\textbf{Baseline Solution.} As our baseline solution, we employ BERT and BERT-RetroMAE without the pretraining method proposed by us. The results are presented in Table \ref{tab:eval}.
+
+\begin{table}[!htb]
+    \centering
+    \begin{tabular}{c|c|c|c|c}
+        & Service Clf & Categorization & Relatedness & Retrieval \\
+        \hline
+        BERT & 0.49 & & & \\
+        BERT-RetroMAE & & & & \\
+        BERT + Contr loss & 0.68 & & &
+    \end{tabular}
+    \caption{Dialogue embedding evaluation results.}
+    \label{tab:eval}
+\end{table}
 
@@
 \appendix
-\section{Dialogue Dataset}
+\section{Dialogue Dataset} \label{app:dataset}
 Large dataset is important for contrastive learning.
@@
     \end{tabular}
-    \caption{All the datasets are taken from DialogStudio collection \cite{zhang2023dialogstudio}}
+    \caption{All the datasets are taken from DialogStudio collection \cite{zhang2023dialogstudio}. BERT tokenizer is used to count \# tokens column.}
     \label{tab:dataset}
@@
         \includegraphics[width=0.9\linewidth]{figures/pairwise.drawio.pdf}
-        \caption{for measuring similarities between utterances}
+        \caption{for measuring similarity between utterances}
         \label{fig:pairwise}
@@
 In a dialogue, it is crucial to compare not just pairs of utterances but pairs of context-response.
-Therefore, we employed the following model (fig.
+Therefore, we employed the following model (Fig.
 \ref{fig:pairwise}):
@@
 
-As a result, the obtained model is capable of recognizing individual stages in dialogues (fig.
+As a result, the obtained model is capable of recognizing individual stages in dialogues (Fig.
 \ref{fig:pairwise-clister-heatmap}).
@@
 
-We trained the special model to merge utterances of two different dialogues (fig.
+We trained the special model to merge utterances of two different dialogues (Fig.
 \ref{fig:listwise}).
@@
 In order to maximize diversity of training data, we use not only 5 basic augmentations described in section \ref{sec:aug}, but also 4 extra compositions of augmentations.
-All the resulting pipelines are defined in fig.
+All the resulting pipelines are defined in Fig.
 \ref{fig:aug-compositions}.

### 2023-11-29 paper/main.tex
added conclusion and reorganized exps section

@@
 % Update your Headers here
-\fancyhead[LO]{Running Title for Header}
+% \fancyhead[LO]{Running Title for Header}
 % \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}
@@
 %% Title
-\title{Contrastive Learning with Intent-Preserving and Intent-Corrupting Augmentations for Training Dialogue Embeddings.
+\title{Contrastive Learning with Augmentations for Training Dialogue Embeddings.
 %%%% Cite as
@@
 
-\texttt{this section is incomplete, this is still a draft }
-
 \subsection{Text Embedding}
 
-One of the breakthroughs in text embedding came from the need for solving the task of semantic textual similarity \cite{reimers2019sentencebert}, i.e.
-comparing texts with each other.
-Previously existing methods were either weak (averaging word embeddings) or required enormous amount of computation (BERT cross-encoder).
-Comparing texts with bi-encoder is performed via averaging hidden states from the last transformer block of language model or taking hidden state of special token \texttt{[CLS]}.
-This setting opens up a possibility for clustering, retrieval, pair classification etc.
-within text domain \cite{muennighoff2023mteb}.
+One of the breakthroughs in text embedding arose from the necessity to address the task of semantic textual similarity \cite{reimers2019sentencebert}, which involves comparing texts.
+Previously existing methods were either weak, such as averaging word embeddings, or demanded a significant computational burden, as with the BERT cross-encoder.
+Text comparison using bi-encoders is accomplished by averaging hidden states from the last transformer block of the language model or by extracting the hidden state of the special token \texttt{[CLS]}.
+This configuration opens up possibilities for tasks such as clustering, retrieval, pair classification, and more within the realm of textual data \cite{muennighoff2023mteb}.
 
@@
 
-The most efficient way of mining positive pairs is to use supervision.
-For example, datasets on NLI (natural language inference), QA (question answering), fact verification, paraphrases contain pairs of semantically similar texts and are appropriate for training text embedding \cite{li2023general}.
-This data is extremely clean, but scarce.
-Another way is to scrap data from web pages, such as QA forums (Quora), social media (Reddit), etc.
-This data is large, but can be noisy.
-When there are no any facility like supervised or scrapped pairs, augmentation is the last variant.
-Two augmented views of the same object must preserve semantics, that are essential for this object, but provide sufficiently hard task for a model of discriminating one positive sample from the rest of negative ones.
+It is most effective to train embeddings using supervised positive pairs.
+For instance, datasets in the domains of natural language inference (NLI), question answering (QA), fact verification, and paraphrases contain pairs of semantically similar texts, making them suitable for training text embeddings \cite{li2023general}.
+This data is exceptionally clean but limited in quantity.
+Another approach involves scraping data from web pages, such as QA forums (e.g., Quora) and social media platforms (e.g., Reddit).
+This data is abundant, but can be noisy.
+When there are no readily available resources like supervised or scraped pairs, augmentation remains a viable option.
+In this case, two augmented views of the same object should preserve the essential semantics of the object while posing a challenging task for a model to discriminate one positive sample from the rest of the negative ones.
 
-Present-day SoTA text embeddings such as BGE \cite{xiao2023cpack}, GTE \cite{li2023general}, E5 \cite{wang2022text} follow one pipeline of training.
+Present-day SoTA text embeddings such as BGE \cite{xiao2023cpack}, GTE \cite{li2023general}, E5 \cite{wang2022text} follow the same pipeline of training.
 First, a retrieval-oriented language model is trained \cite{xiao2022retromae, gao-callan-2021-condenser}, that is a BERT-like model that can efficiently aggregate global information about input sequence.
-Second, a general-purpose fine-tuning is performed on a large weakly supervised text pairs with large batch contrastive learning.
-Third, task-specific fine-tuning is performed on supervised text pairs with
-contrastive learning with hard negative samples.
+Second, a general-purpose fine-tuning process is conducted using a large corpus of weakly supervised text pairs with large-batch contrastive learning.
+Finally, task-specific fine-tuning on supervised text pairs occurs employing contrastive learning with hard negative samples.
 
@@
 
-AugSBERT \cite{thakur2021augmented} cannot be viewed as the similar approach to ours. Its augmentation uses an already presented dataset of text pairs, which is absent in our case.
+\textbf{Text embedding.} Several existing text embedding methods employ contrastive learning with augmentation-mined positive and negative pairs.
+CERT \cite{fang2020cert} utilizes back-translation to mine positive pairs, while ConSERT \cite{yan-etal-2021-consert} employs various augmentations at the token level.
+Doc2vecC \cite{luo2021unsupervised} relies on thesaurus-based augmentations and back-translation for its approach.
 
-\textbf{Text embedding.} CERT \cite{fang2020cert} utilizes back-translation for mining positive pairs.
-ConSERT \cite{yan-etal-2021-consert} utilizes various augmentations at the token-level.
-Doc2vecC \cite{luo2021unsupervised} utilizes thesaurus-based augmentations and back-translation.
+\textbf{Dialogue embedding.} Dial2vec \cite{liu2022dial2vec} and DialogueCSE \cite{liu2021dialoguecse} modify transformer architecture by adding cross-attention between different groups of utterances after all transformer layers.
+To get a positive pair, both works employ the idea of self-guidence that is similar to ours, because they get two views of the same object.
+In Dial2vec, it's local vs.
+global information about a dialogue, whereas in DialogueCSE, it's the utterances of the first interlocutor vs.
+the utterances of the second interlocutor.
 
-\textbf{Dialogue embedding.} Dial2vec \cite{liu2022dial2vec} modifies transformer architecture by adding cross-attention between utterances of different participants.
-DialogueCSE \cite{liu2021dialoguecse} adds cross-attention between utterances of different stages of a dialogue.
+Our approach stands apart from the methods mentioned above for the following reasons:
+\begin{itemize}
+    \item Our set of augmentations is more potent than the one employed in CERT, comprising a comprehensive set of transformations that facilitate the creation of non-obvious examples, beyond mere paraphrasing.
+    \item Our augmentations adhere to the set of valid objects, preserving grammatical and meaningful constructs.
+They do not disrupt the structure of the dialogue, as they are meticulously designed with context-aware language models, unlike ConSERT, which employs simple token shuffling and deletion, retaining only lexical information.
+    \item our approach does not modify BERT architecture
+\end{itemize}
 
-Our approach is distinct from these above:
-\begin{itemize}
-    \item our augmentations do not take objects outside the set of valid objects (do not violate grammars and meaningfulness), because our augmentations are thoroughly designed with context-aware language models in contrast to ConSERT that simply shuffles and deletes tokens leaving only lexical info in touch
-    \item our augmentations form a full-fledged set of transformations that construct non-obvious examples, not just paraphrasing or token-level replacements
-    \item our approach doesn't modify BERT architecture
-\end{itemize}
+TODO: add text about SimCLR?
 
@@
 
-\subsection{Dialogue Encoder Architecture}
+\subsection{Encoder Fine-tuning}
 
-As a model for embedding, we try BERT \cite{devlin2019bert}, RoBERTa \cite{liu2019roberta}, RetroMAE \cite{xiao2022retromae} without any modifications.
+As a model for embedding, we try BERT-like models without any modifications \cite{devlin2019bert}.
 The input is \texttt{[CLS] ut1 [SEP] ut2 [SEP] ut3}, and the output is the hidden state of \texttt{[CLS]} token from the last layer.
@@
 
-\subsection{Pre-train Task} \label{sec:pretrain}
+In our fine-tuning, a positive pair is obtained with intent-preserving augmentations of the same dialogue, and the remaining dialogues from the training batch are used as negative samples.
+In order to reduce required batch size, we utilize intent-corrupting augmentations as hard negative samples.
 
-In our case, $a, b$ are two intent-preserving augmentations of the same dialogue, and the remaining dialogues from the training batch are used as negative examples.
-This task simulates the second stage of training general-purpose text embeddings, which uses a large batch and weakly supervised text pairs.
-In addition to this, in order to reduce required batch size, we utilize intent-corrupting augmentations as hard negative samples.
-Notably, the technique with two augmentations and dual encoder resembles BYOL \cite{grill2020bootstrap}, well-known method for self-supervised learning \texttt{(maybe move this sentence to Related Works)}.
-
-\subsection{Evaluation}
+\subsection{Evaluation} \label{sec:evaluation}
 
 We evaluate dialogue embeddings in transfer learning setting.
-Specifically, our evaluation methods utilize embeddings as features, without modifying the encoder model itself.
+Specifically, our evaluation methods utilize frozen embeddings as features.
 Inspired by \cite{liu2022dial2vec} we employ these evaluation methods: 1) domain classification, 2) dialogue retrieval, 3) intent similarity.
@@
 F1-macro is used to measure quality.
+This evaluation method can demonstrate how well embeddings encode some implicit features about the dialogue and its participants in perspective.
 
@@
 Mean average precision at 100 is used to measure quality.
+This evaluation method can demonstrate potential effectiveness for retrieval and other downstream tasks like clustering.
 
 \textbf{Intent similarity.} We sample pairs of dialogues from train and validation splits of dataset.
-A linear classifier is trained on the former pairs and evaluated on the latter pairs using Pearson correlation between predicted similarity scores and gold ones.
+A linear regression is trained on the former pairs and evaluated on the latter pairs using Pearson correlation between predicted similarity scores and gold ones.
 The gold scores are obtained using DGAC clustering \cite{dgac}.
 Clusters represent intents of dialogue participants.
-Hence, each dialogue can be associated with a set of intents.
+Each dialogue can be associated with a set of intents.
 We define gold intent similarity of two dialogues as a dice similarity between their sets of intents.
@@
 
-\textbf{Dataset.} In all experiments, we used the same dataset of dialogues.
-This dataset is a combination of several task-oriented dialogue datasets from DialogStudio \cite{zhang2023dialogstudio}.
+this section is incomplete
+
+\textbf{Dataset.} Large dataset is important for contrastive learning.
+In all experiments, we used the same dataset of dialogues.
+This dataset is a combination of several task-oriented dialogue datasets from DialogStudio \cite{zhang2023dialogstudio} which are listed in Table \ref{tab:dataset}.
 All the dialogues were filtered based on their length, resulting in a dataset comprising 501K dialogues.
-For more details on the dataset used, please refer to Appendix \ref{app:dataset}.
-
-\textbf{Baseline Solution.} As our baseline solution, we employ BERT and BERT-RetroMAE without the pretraining method proposed by us. The results are presented in Table \ref{tab:eval}.
-
-\begin{table}[!htb]
-    \centering
-    \begin{tabular}{c|c|c|c|c}
-        & Service Clf & Categorization & Relatedness & Retrieval \\
-        \hline
-        BERT & 0.49 & & & \\
-        BERT-RetroMAE & & & & \\
-        BERT + Contr loss & 0.68 & & &
-    \end{tabular}
-    \caption{Dialogue embedding evaluation results.}
-    \label{tab:eval}
-\end{table}
-
-\section*{Acknowledgments}
-This was was supported in part by......
-
-%Bibliography
-\bibliographystyle{unsrt}  
-\bibliography{references}
-
-\appendix
-\section{Dialogue Dataset} \label{app:dataset}
-Large dataset is important for contrastive learning.
-For training our models, we took a merge of some task-oriented datasets from DialogStudio collection \cite{zhang2023dialogstudio}.
-All of them are listed in the table \ref{tab:dataset}.
 
@@
 \end{table}
+
+First, we evaluate pre-trained models BERT \cite{devlin2019bert}, RoBERTa \cite{liu2019roberta}, RetroMAE \cite{xiao2022retromae}.
+Results are presented in Table \ref{tab:eval-raw}.
+This table describes existing results, that can be obtained by available model weights without any fine-tuning on dialogue data.
+
+\begin{table}[!htb]
+    \centering
+\begin{tabular}{l|c|c|c}
+         & Classification & Retrieval & Similarity \\ \hline
+BERT     & $47.53$          & $66.31$     &            \\
+RoBERTa  & $48.66$          & $84.40$     &            \\
+RetroMAE & $69.84$          & $86.95$     &            \\
+\end{tabular}
+    \caption{Evaluation of available models. For the meaning of these values, see Section \ref{sec:evaluation}.}
+    \label{tab:eval-raw}
+\end{table}
+
+\textbf{(Weak) Training.} In our fine-tuning approach, we used the following settings: a batch size of 32, AdamW optimizer with a weight decay of 1e-2, and a fixed learning rate of 3e-6.
+We froze all transformer layers except the last one.
+It's worth noting that we applied only a partial set of augmentations and worked with just 10\% of the entire dataset.
+Evaluation results after one epoch of our contrastive fine-tuning on dialogue data are presented in Table \ref{tab:eval-one-epoch}.
+
+\begin{table}[!htb]
+    \centering
+\begin{tabular}{l|c|c|c}
+         & Classification & Retrieval      & Similarity \\ \hline
+BERT     & $67.18$ $(+19.65)$ & $77.99$ $(+11.68)$ &            \\
+RoBERTa  & $68.20$ $(+19.54)$ & $79.94$ $(-4.46)$  &            \\
+RetroMAE & $70.33$ $(+0.49)$  & $89.08$ $(+2.13)$  &           
+\end{tabular}
+    \caption{Evaluation after one epoch of contrastive fine-tuning.
+Number in parentheses is a difference with results in Table \ref{tab:eval-raw}.
+For the meaning of these values, see Section \ref{sec:evaluation}.}
+    \label{tab:eval-one-epoch}
+\end{table}
+
+\section{Limitations.}
+
+Out method has several limitations.
+
+\begin{itemize}
+    \item The weights obtained are specific to the language used.
+Consequently, it is necessary to gather dialogues in the target language, train auxiliary pairwise and listwise models, and subsequently employ contrastive fine-tuning on a BERT-like encoder.
+An alternative approach to address this challenge is to accumulate multilingual data.
+    \item All experiments were conducted with a complete dialogue inputted directly into the transformer, which is constrained by the maximum number of tokens in the input sequence.
+One potential resolution is to utilize the HSSA model, designed for handling individual utterances.
+However, this option may result in a reduction in quality.
+\end{itemize}
+
+\section{Conclusion.}
+
+A study has been conducted on building embeddings for task-oriented dialogues.
+A procedure for contrastive learning with augmentations that preserve and corrupt interlocutors' intents was proposed.
+The effectiveness of this procedure was demonstrated for BERT-like models in tasks such as dialogue classification, retrieval, and similarity.
+An ablation study was conducted, showing the contribution of each type of augmentation to the final embeddings.
+
+%Bibliography
+\bibliographystyle{unsrt}  
+\bibliography{references}
+
+\appendix
 
